---
title: "GSEA Analysis"
author: "Saumya"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_depth: 3
    toc_float: true
    theme: cosmo
#    number_sections: true #adds numbers to headers
    code_folding: hide
---


```{r, results = FALSE, message = FALSE}
# Load required libraries
library(openxlsx)
library(clusterProfiler)
library(msigdbr)
library(org.Hs.eg.db)
library(tidyverse)
library(readxl)
library(rmarkdown)
library(ggplot2)
library(fgsea)
library(msigdbdf)
```

<!-- Information of Used Library -->
## Information of Used R Libraries

The following R libraries were used to process the data and produce this report.


R: `r R.Version()$version.string`,

openxlsx: `r packageVersion('openxlsx')`,

clusterProfiler: `r packageVersion('clusterProfiler')`, 

msigdbr: `r packageVersion('msigdbr')`, 

org.Hs.eg.db: `r packageVersion('org.Hs.eg.db')`, 

tidyverse: `r packageVersion('tidyverse')`,

readxl: `r packageVersion('readxl')`, 

rmarkdown: `r packageVersion('rmarkdown')`,

ggplot2: `r packageVersion('ggplot2')`

<!-- Run GSEA (Clusterprofiler) for only one input csv -->

```{r, results = FALSE,  message=FALSE, warning=FALSE}
# Prepare gene sets for human

set.seed(2024)
human_gobp_sets <- msigdbr(species = "Homo sapiens", category = "C5", subcategory = "GO:BP")
human_kegg_sets <- msigdbr(species = "Homo sapiens", category = "C2", subcategory = "CP:KEGG_LEGACY")
human_reactome_sets <- msigdbr(species = "Homo sapiens", category = "C2", subcategory = "CP:REACTOME")

human_gene_sets <- rbind(human_gobp_sets, human_kegg_sets, human_reactome_sets)
```

```{r, results=FALSE, message=FALSE, warning=FALSE}
# Define the GSEA function for human

gsea_human <- function(df, condition){
  df_cluster <- df %>% dplyr::filter(cluster == condition)
  lfc_vector <- df_cluster$avg_log2FC
  names(lfc_vector) <- df_cluster$gene
  lfc_vector <- sort(lfc_vector, decreasing = TRUE)

  gsea_results <- GSEA(
    geneList = lfc_vector, # Ordered ranked gene list
    minGSSize = 5, # Minimum gene set size
    maxGSSize = 500, # Maximum gene set size
    eps = 0, # Boundary for calculating the p value
    seed = TRUE, # Set seed to make results reproducible
    pvalueCutoff = 1,
    pAdjustMethod = "BH", # Benjamini-Hochberg correction
    TERM2GENE = dplyr::select(human_gene_sets, gs_name, gene_symbol)
  )
  return(gsea_results@result)
}
```

```{r, eval=FALSE, results = FALSE, message=FALSE, warning=FALSE}
# Read the input CSV file
degs_samples <- read_csv("../csvs/data_all_markers_condition.csv")

# Define conditions (clusters) from input csv file
conditions_samples <- as.character(unique(degs_samples$cluster))
```

```{r, eval=FALSE, results = FALSE, message=FALSE, warning=FALSE}
# Perform GSEA for all conditions and collect results

gsea_results_samples<- data.frame()

 for (i in conditions_samples) {
  gsea_samples_tmp <- gsea_human(degs_samples, i) %>% dplyr::mutate(cluster = i)
  gsea_results_samples <- rbind(gsea_results_samples, gsea_samples_tmp)
}
```

```{r, eval=FALSE, results = FALSE, message=FALSE, warning=FALSE}
# Split the results by condition

results_list_samples <- list()

for (type in conditions_samples) {
  type_results_samples <- gsea_results_samples %>% dplyr::filter(cluster == type) %>% dplyr::select(-cluster)
  results_list_samples[[type]] <- type_results_samples
 }
```

```{r, eval=FALSE, results = FALSE, message=FALSE, warning=FALSE}
# Create a list of data frames for writing to Excel. While making the barplot by using this file, the results would have to be subsetted for only one condition

write.xlsx(gsea_results_samples, file = "../csvs/GSEA_Clusterprofiler_Condition.xlsx")
```



<!-- Run GSEA (Clusterprofiler) for multiple input csv  (like subclusters) -->


```{r, results = FALSE, eval = FALSE, message=FALSE, warning=FALSE}
# Code block to find condition wise markers for each subcluster of subsetted data. Same code as present in 04_Subset_Clustering.Rmd file

# Read Seurat object
subset_data <- readRDS("../RDS/subset_umap_10_clusters.RDS")

# Ensure RNA assay is set
DefaultAssay(subset_data) <- "RNA"
Idents(subset_data) <- "seurat_clusters"

# Store subcluster identities (backup)
subset_data$subcluster <- Idents(subset_data)

# Set condition as active identity
Idents(subset_data) <- "condition"

subset_data <- JoinLayers(subset_data)

# Create an empty list to store results
markers_list <- list()

# Loop over each subcluster
for (cluster in unique(subset_data$subcluster)) {
  # Subset data for the current subcluster
   message("Processing Cluster: ", cluster)
   subcluster_data <- subset(subset_data, subset = subcluster == cluster)
 
  # Set condition as active identity
  Idents(subcluster_data) <- "condition"

  # Find all markers between Condition and Control within this subcluster
  markers <- FindAllMarkers(subcluster_data, only.pos = FALSE, min.pct = 0.25, logfc.threshold = 0.25)
   markers$p_val_adj <- p.adjust(markers$p_val, method = "fdr")

  # Store results with subcluster name
   markers_list[[paste0("subset_subcluster_", cluster)]] <- markers
}

# Save results to CSV
for (name in names(markers_list)) {
  write.csv(markers_list[[name]], file=paste0("../csvs/subset/condition/", name, "_Condition_vs_Control.csv"))
}
```


```{r, eval=FALSE, results = FALSE, message = FALSE}
# Set the path to your CSV files
csv_dir <- "../csvs/condition/"

# Prepare gene sets for human
# Picking the same categories as used for DAVID analysis
set.seed(2024)
human_gobp_sets <- msigdbr(species = "Homo sapiens", category = "C5", subcollection = "GO:BP")
human_kegg_sets <- msigdbr(species = "Homo sapiens", category = "C2", subcollection = "CP:KEGG_LEGACY")
human_reactome_sets <- msigdbr(species = "Homo sapiens", category = "C2", subcollection = "CP:REACTOME")

human_gene_sets <- rbind(human_gobp_sets, human_kegg_sets, human_reactome_sets)
```

```{r, results=FALSE, message=FALSE, warning=FALSE}
# Define the GSEA function for human

gsea_human <- function(df, condition){
  df_cluster <- df %>% dplyr::filter(cluster == condition)
  lfc_vector <- df_cluster$avg_log2FC
  names(lfc_vector) <- df_cluster$gene
  lfc_vector <- sort(lfc_vector, decreasing = TRUE)

  gsea_results <- GSEA(
    geneList = lfc_vector, # Ordered ranked gene list
    minGSSize = 5, # Minimum gene set size
    maxGSSize = 500, # Maximum gene set size
    eps = 0, # Boundary for calculating the p value
    seed = TRUE, # Set seed to make results reproducible
    pvalueCutoff = 1,
    pAdjustMethod = "BH", # Benjamini-Hochberg correction
    TERM2GENE = dplyr::select(human_gene_sets, gs_name, gene_symbol)
  )
  return(gsea_results@result)
}
```

```{r, eval=FALSE, results = FALSE, message = FALSE}
# List all relevant CSV files
csv_files <- list.files(csv_dir, pattern = "_Condition_vs_Control.csv", full.names = TRUE)

# Create a master list to collect results
all_gsea_results <- list()

# Loop through each file
for (csv_path in csv_files) {
  subcluster_name <- tools::file_path_sans_ext(basename(csv_path))
  message("Processing: ", subcluster_name)

  df <- read.csv(csv_path)

  # Skip subclusters with insufficient gene info
  if (!("gene" %in% colnames(df)) || nrow(df) < 10) {
    warning("Skipping ", subcluster_name, " due to missing 'gene' column or low gene count.")
    next
  }
  condition <- "Condition"
  gsea_result <- gsea_human(df, condition)


  # Append results with an identifying column
  gsea_result$subcluster <- subcluster_name
  all_gsea_results[[subcluster_name]] <- gsea_result
}

# Combine and export
combined_results <- do.call(rbind, all_gsea_results)
write.xlsx(combined_results, file = "../csvs/condition/combined_GSEA_results.xlsx")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Read the sheets into data frames. If reading in single csv file, then subset data to only condition of interest
samples_results <- read_excel("../csvs/condition/combined_GSEA_results.xlsx")
```

## ClusterProfiler Results {.tabset}

### Understanding the output

The output of the enrichment analysis is an excel file divided into subsheets based on clusters (JAK2 conditions) which has 11 headers and explanation on each of the headers can be found below:

1. ID: Identifier for the gene set. This is the identifier of the pathway or gene ontology term associated with the gene set.

2. Description: Description of the gene set. This typically provides additional details about the pathway or biological process represented by the gene set.

3. setSize: Size of the gene set, i.e., the number of genes included in the gene set.

4. enrichmentScore: The enrichment score computed by the gene set enrichment analysis. This score indicates the degree to which the gene set is overrepresented at the top or bottom of a ranked list of genes (based on fold changes or another ranking metric).  A higher enrichmentScore (closer to 1 or -1) indicates stronger enrichment, which is desirable as it suggests that the gene set is significantly overrepresented in the analyzed dataset. 

5. NES (Normalized Enrichment Score): The normalized enrichment score, which is the enrichment score normalized to account for differences in gene set sizes and correlations between gene sets and the data being analyzed. Higher NES is desired and considered more reliable whereas A NES close to 0 indicates no enrichment.

6. pvalue: The p-value associated with the enrichment score. It indicates the statistical significance of the enrichment of the gene set in the ranked list of genes. Lower p value is better.

7. p.adjust: The adjusted p-value after multiple testing correction. Adjusted using Benjamini-Hochberg (BH) procedure. Lower values are better.

8. qvalue: The q-value, which is another measure of significance adjusted for multiple testing. It is similar to the adjusted p-value. Lower values are better.

9. rank: The rank at which the gene set achieves its maximum enrichment score in the ranked list of genes. A lower rank indicates that the gene set is enriched towards the top of the ranked list of genes, which is desirable as it suggests higher significance. 

10. leading_edge: Information about the leading edge subset of genes contributing to the enrichment score. This typically includes the percentage of tags, list, and signal genes in the leading edge. Higher percentage contribution indicates a substantial contribution of these genes to the observed enrichment.

11. core_enrichment: Specific genes involved in the core enrichment of the gene set. These are often the key members of the gene set contributing to its enrichment score.

```{r,results='asis', message=FALSE, warning=FALSE}
# Display the contents of the sheets using paged_table
cat('###', "Samples Results", ' \n')
paged_table(samples_results)
cat(' \n \n')
```

```{r, results=FALSE, message=FALSE, warning=FALSE}
# Function to plot up/down pathways 
plot_top_pathways <- function(gsea_results, top_n = 10) {
  
  significant_results <- gsea_results %>%
    dplyr::filter(p.adjust < 0.05)
  
  top_up <- significant_results %>%
    filter(NES > 0) %>%
    arrange(desc(NES)) %>%
    head(top_n)
  
  top_down <- significant_results %>%
    filter(NES < 0) %>%
    arrange(NES) %>%
    head(top_n)
  
  up_plot <- if (nrow(top_up) > 0) {
    ggplot(top_up, aes(x = reorder(Description, NES), y = NES)) +
      geom_bar(stat = "identity", fill = "darkred") +
      coord_flip() +
      labs(title = "Top 10 Upregulated Pathways", x = "Pathway", y = "NES") +
      theme_minimal() +
      theme(axis.text.y = element_text(size = 6))
  } else {
    ggplot() + 
      annotate("text", x = 1, y = 1, label = "No significantly upregulated pathways") + 
      theme_void()
  }

  down_plot <- if (nrow(top_down) > 0) {
    ggplot(top_down, aes(x = reorder(Description, -NES), y = NES)) +
    geom_bar(stat = "identity", fill = "skyblue") +
    coord_flip() +
    labs(title = "Top 10 Downregulated Pathways Based on NES", x = "Pathway", y = "Normalized Enrichment Score (NES)") +
    theme_minimal() +
    theme(axis.text.y = element_text(size = 6))
  } else {
    ggplot() + 
      annotate("text", x = 1, y = 1, label = "No significantly downregulated pathways") + 
      theme_void()
  }

  return(list(up_plot = up_plot, down_plot = down_plot))
}
```


```{r, results = FALSE, message = FALSE, warning=FALSE}
# Store plots by cluster

plots_by_cluster <- list()
unique_clusters <- unique(samples_results$subcluster)

for (clust in unique_clusters) {
  clust_data <- samples_results %>% filter(subcluster == clust)
  plots_by_cluster[[as.character(clust)]] <- plot_top_pathways(clust_data)
}
```

## Barplots {.tabset}

Top 10 Upregulated and Downregulated Pathways based on NES

```{r,results='asis', message=FALSE, warning=FALSE,fig.width=10,fig.height=6}
# Change gsub from cluster to subcluster if required based on file name

for (clust in names(plots_by_cluster)) {
  subcluster_clean <- gsub(".*?(cluster_\\d+).*", "\\1", clust)

  cat('###', subcluster_clean, '{.tabset}\n\n')
  
  cat('#### Upregulated\n\n')
  print(plots_by_cluster[[clust]]$up_plot)
  cat('\n\n')
  
  cat('#### Downregulated\n\n')
  print(plots_by_cluster[[clust]]$down_plot)
  cat('\n\n')
}

```

