---
title: "Sample Integration"
author: "Saumya"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_depth: 3
    toc_float: true
    theme: cosmo
#    number_sections: true # Adds numbers to headers
    css: styles.css
    code_folding: hide
---

```{r, results = FALSE, message=FALSE, warning=FALSE}
# Load packages
library(dplyr)
library(Seurat)
library(patchwork)
library(ggplot2)
library(cowplot)
library(rmarkdown)
library(SummarizedExperiment)
library(htmlwidgets)
library(htmltools)
library(shiny)
library(openxlsx)
library(scDblFinder)
library(hdf5r)
library(harmony)
```

<!-- Information of Used Library -->
## Information of Used R Libraries

The following R libraries were used to process the data and produce this report.

R: `r R.Version()$version.string`,

dplyr: `r packageVersion('dplyr')`, 

Seurat: `r packageVersion('Seurat')`, 

patchwork: `r packageVersion('patchwork')`, 

ggplot2: `r packageVersion('ggplot2')`, 

cowplot: `r packageVersion('cowplot')`,

scDblFinder: `r packageVersion('scDblFinder')`

```{r, results = FALSE, message=FALSE, warning=FALSE}
sampleNames <- c("ABC", "XYZ", "LMN")
data_dir <- ("../../raw_h5")
projectName <- "Seurat"
h5_string <- "_filtered_feature_bc_matrix.h5"
```

## Integration of data to remove artifacts/batch effect

```{r,  eval=FALSE, message = FALSE, results= FALSE, warning=FALSE}
# Use Harmony for Integration

data.combined <- readRDS("../RDS/data_dblt_filtered_umap.RDS")

data.combined <- RunHarmony(
  object = data.combined,
  group.by.vars = "sample_names",  # Adjust based on what variable it needs to be integrated on
  theta = c(2),             # Reduce batch correction strength. Based on number of group.by.vars. Default is 2
  lambda = 1,             # Higher the number, softer the batch correction. Default = 1
  dims.use = 1:20,          # Can reduce dims. We have only 20 that we chose according to the ElbowPlot
)

# Save data
saveRDS(data.combined, file = paste("../RDS/data_Integrated.RDS") )
```

```{r,  eval=FALSE, message = FALSE, results= FALSE, warning=FALSE}
# Load data
data.combined <- readRDS("../RDS/data_Integrated.RDS")
```


```{r,  eval=FALSE, message = FALSE, results= FALSE, warning=FALSE}
# Run the standard workflow for visualization and clustering
data.combined <- ScaleData(data.combined, verbose = FALSE)

# Run UMAP using the Harmony embeddings, not the original PCA
data.combined <- RunUMAP(data.combined, reduction = "harmony", dims = 1:20)
 
# Save integrated data
saveRDS(data.combined, file = paste("../RDS/data_Integrated_umap.RDS") )
```


# UMAP for Samples before integration {.tabset}

```{r,results='asis', message=FALSE, warning=FALSE}
# Load non-integrated data
data.bf <- readRDS("../RDS/data_combined_raw_umap.RDS")

cat('##','All samples',' \n')
DimPlot(data.bf, reduction = "umap", group.by= 'sample_names' ) + ggtitle('Samples before integration')+ theme(legend.position = "none")
cat(' \n \n')

# Sample wise dimplot
for (samp in sample_names_list) {
  cat('##',samp,' \n')
  print(DimPlot(subset(data.bf, subset = (sample_names %in% c(samp))), reduction = "umap") + ggtitle(samp))
 cat(' \n \n')
}
```

# UMAP for Samples after integration {.tabset}

```{r,results='asis', message=FALSE, warning=FALSE}
# Load integrated data
data.combined <- readRDS("../RDS/data_Integrated_umap.RDS")

cat('##','All samples',' \n')
DimPlot(data.combined, reduction = "umap", group.by= 'sample_names' ) + ggtitle('Samples after integration') + theme(legend.position = "none") 
cat(' \n \n')

# Sample wise dimplot
for (samp in sample_names_list) {
  cat('##',samp,' \n')
  print(DimPlot(subset(data.combined, subset = (sample_names %in% c(samp))), reduction = "umap") + ggtitle(samp))
 cat(' \n \n')
}
```





