---
title: "Gene enrichment for 3 Oligo Subclusters"
author: "Saumya"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---
  
```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# Markdown parameters
set_echo = FALSE
set_eval = FALSE
```

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# Function to set markdown text color
colorize <- function(x, color) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{%s}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'>%s</span>", color,
            x)
  } else x
}
```

```{r, include=FALSE}
# Set height of chunk output
options(width = 60)
local({
  hook_output <- knitr::knit_hooks$get('output')
  knitr::knit_hooks$set(output = function(x, options) {
    if (!is.null(options$max.height)) options$attr.output <- c(
      options$attr.output,
      sprintf('style="max-height: %s;"', options$max.height)
    )
    hook_output(x, options)
  })
})
```
  
```{r, results = FALSE, echo=set_echo, message=FALSE, warning=FALSE}
# Load packages
library("ggplot2")
library("dplyr")
library('AnnotationDbi')
library("rmarkdown")
#library("RDAVIDWebService")
library("org.Mmu.eg.db")
library("xlsx")
library("readxl")
library("openxlsx")
#library('rrvgo')
library(stringr)
```

<!-- Information of Used Library -->
### `r colorize('Information of Used R Libraries', 'blue')` 

The following R libraries were used to process the data and produce this report.


R: `r R.Version()$version.string`,

dplyr: `r packageVersion('dplyr')`,

RDAVIDWebService: `r packageVersion('RDAVIDWebService')`, 

org.Mmu.eg.db: `r packageVersion('org.Mmu.eg.db')`, 

xlsx: `r packageVersion('xlsx')`,

readxl: `r packageVersion('readxl')`,

rrvgo: `r packageVersion('rrvgo')`.

RDAVIDWebService requires R version 4.0.


```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Useful string constant
suffix <- "Seurat"
sampleNames_filt <- c("RD_VA_01", "RD_VA_02", "RD_VA_03", "RD_VA_05",  
                 "RD_VA_07", "RD_VA_08", "RD_VA_09", "RD_VA_10","RD_VA_11","RD_VA_12")
sample_Names_10 = gsub("_", "-", sampleNames_filt)
rawDataDir_1 <- "/projectnb/czlab/A10_Rosene/Data/20230408_combinedTwo_results/"
rawDataDir_2 <- "/outs/filtered_feature_bc_matrix"
projectName <- "RD_VA_01_12"

# Seurat parameters
variable.features <- 2000
nPCs <- 50
```


```{r, echo=FALSE}
replacENTREZID <- function(functionMatrix, mapping) {
  for(i in 1:nrow(functionMatrix)) {
    genelist <- trimws(strsplit(functionMatrix[i,]$Genes, ",")[[1]])
    geneName <- mapping[match(genelist, mapping$ENTREZID), 1]
    functionMatrix[i,]$Genes <- paste(geneName, collapse = ",")
  }
  return(functionMatrix)
}
```

```{r, echo=FALSE}
num_of_used_genes=500
group_names = c('Mature Oligo', 'Immature Oligo', 'Synaptic Oligo')
```

```{r, echo=FALSE}
#Function to plot results
plotRrvgo <- function(fileName) {
  options(java.parameters = "-Xmx8000m")
  upFunc <- fileName
  upFunc <- upFunc[upFunc$Category=="GOTERM_BP_DIRECT",]
  upFunc$Category <- substring(upFunc$Term,0,10)
  simMatrix <- calculateSimMatrix(upFunc$Category,
                                  orgdb="org.Mmu.eg.db",
                                  ont="BP",
                                  method="Rel")
  scores <- setNames(-log10(as.numeric(upFunc$PValue)), upFunc$Category)
  reducedTerms <- reduceSimMatrix(simMatrix,
                                  scores,
                                  threshold=0.9,
                                  orgdb="org.Mmu.eg.db")
  
  return(list(first=simMatrix, second=reducedTerms))
}

```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#Input positive marker gene list and then take top 500 genes based on condition of interest.

# Input positive marker csv file
positive_markers_mature <- read.csv("data_positive_markers_mature_oligo.csv")
positive_markers_immature <- read.csv("data_positive_markers_immature_oligo.csv")
positive_markers_synaptic <- read.csv("data_positive_markers_synaptic_oligo.csv")
positive_markers <- read.csv("data_positive_markers_oligo.csv")

# Split the table into two based on cluster and select top 200 genes for each

markers_CR <- positive_markers %>%
  filter(cluster == "CR") %>%
  slice_min(order_by = p_val, n = 200) %>%
  dplyr::select(gene)

markers_CONTROL <- positive_markers %>%
  filter(cluster == "CONTROL") %>%
  slice_min(order_by = p_val, n = 200) %>%
  dplyr::select(gene)

mature_markers_CR <- positive_markers_mature %>%
  filter(cluster == "CR") %>%
  slice_min(order_by = p_val, n = 200) %>%
  dplyr::select(gene)

mature_markers_CONTROL <- positive_markers_mature %>%
  filter(cluster == "CONTROL") %>%
  slice_min(order_by = p_val, n = 200) %>%
  dplyr::select(gene)

immature_markers_CR <- positive_markers_immature %>%
  filter(cluster == "CR") %>%
  slice_min(order_by = p_val, n = 200) %>%
  dplyr::select(gene)

immature_markers_CONTROL <- positive_markers_immature %>%
  filter(cluster == "CONTROL") %>%
  slice_min(order_by = p_val, n = 200) %>%
  dplyr::select(gene)

synaptic_markers_CR <- positive_markers_synaptic %>%
  filter(cluster == "CR") %>%
  slice_min(order_by = p_val, n = 200) %>%
  dplyr::select(gene)

synaptic_markers_CONTROL <- positive_markers_synaptic %>%
  filter(cluster == "CONTROL") %>%
  slice_min(order_by = p_val, n = 200) %>%
  dplyr::select(gene)

# Create a new workbook for oligo
wb <- createWorkbook()

# Add sheets
addWorksheet(wb, "Oligo_CR")
addWorksheet(wb, "Oligo_CONTROL")

# Write data to sheets
writeData(wb, sheet = "Oligo_CR", x = markers_CR)
writeData(wb, sheet = "Oligo_CONTROL", x = markers_CONTROL)

# Save the workbook
saveWorkbook(wb, "oligo_genelist_for_DAVID.xlsx", overwrite = TRUE)



# Create a new workbook for mature oligo
wb_mature <- createWorkbook()

# Add sheets
addWorksheet(wb_mature, "Mature_Oligo_CR")
addWorksheet(wb_mature, "Mature_Oligo_CONTROL")

# Write data to sheets
writeData(wb_mature, sheet = "Mature_Oligo_CR", x = mature_markers_CR)
writeData(wb_mature, sheet = "Mature_Oligo_CONTROL", x = mature_markers_CONTROL)

# Save the workbook
saveWorkbook(wb_mature, "mature_oligo_genelist_for_DAVID.xlsx", overwrite = TRUE)

# Create a new workbook for immature oligo
wb_immature <- createWorkbook()

# Add sheets
addWorksheet(wb_immature, "Immature_Oligo_CR")
addWorksheet(wb_immature, "Immature_Oligo_CONTROL")

# Write data to sheets
writeData(wb_immature, sheet = "Immature_Oligo_CR", x = immature_markers_CR)
writeData(wb_immature, sheet = "Immature_Oligo_CONTROL", x = immature_markers_CONTROL)

# Save the workbook
saveWorkbook(wb_immature, "immature_oligo_genelist_for_DAVID.xlsx", overwrite = TRUE)

# Create a new workbook for synaptic oligo
wb_synaptic <- createWorkbook()

# Add sheets
addWorksheet(wb_synaptic, "Synaptic_Oligo_CR")
addWorksheet(wb_synaptic, "Synaptic_Oligo_CONTROL")

# Write data to sheets
writeData(wb_synaptic, sheet = "Synaptic_Oligo_CR", x = synaptic_markers_CR)
writeData(wb_synaptic, sheet = "Synaptic_Oligo_CONTROL", x = synaptic_markers_CONTROL)

# Save the workbook
saveWorkbook(wb_synaptic, "synaptic_oligo_genelist_for_DAVID.xlsx", overwrite = TRUE)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#Read in Oligo subcluster gene lists which have genes based on p_value < 0.01

genelist_CR <- read.xlsx("oligo_genelist_for_DAVID.xlsx", sheet= "Oligo_CR")
genelist_CONTROL <- read.xlsx("oligo_genelist_for_DAVID.xlsx", sheet= "Oligo_CONTROL")
mature_genelist_CR <- read.xlsx("mature_oligo_genelist_for_DAVID.xlsx", sheet= "Mature_Oligo_CR")
mature_genelist_CONTROL <- read.xlsx("mature_oligo_genelist_for_DAVID.xlsx", sheet= "Mature_Oligo_CONTROL")
immature_genelist_CR <- read.xlsx("immature_oligo_genelist_for_DAVID.xlsx", sheet= "Immature_Oligo_CR")
immature_genelist_CONTROL <- read.xlsx("immature_oligo_genelist_for_DAVID.xlsx", sheet= "Immature_Oligo_CONTROL")
synaptic_genelist_CR <- read.xlsx("synaptic_oligo_genelist_for_DAVID.xlsx", sheet= "Synaptic_Oligo_CR")
synaptic_genelist_CONTROL <- read.xlsx("synaptic_oligo_genelist_for_DAVID.xlsx", sheet= "Synaptic_Oligo_CONTROL")
```

## `r colorize('Input gene lists containing 200 genes for 3 Oligo Subclusters sorted by p value', 'blue')` {.tabset}

```{r,results='asis', echo=FALSE, message=FALSE, warning=FALSE}
cat('###', "Oligo CR", ' \n')
paged_table(genelist_CR)
cat('\n \n')
cat('###', "Oligo CONTROL", ' \n')
paged_table(genelist_CONTROL)
cat('\n \n')
```


```{r,results='asis', echo=FALSE, message=FALSE, warning=FALSE}
cat('###', "Mature Oligo CR", ' \n')
paged_table(mature_genelist_CR)
cat('\n \n')
cat('###', "Mature Oligo CONTROL", ' \n')
paged_table(mature_genelist_CONTROL)
cat('\n \n')
```

```{r,results='asis', echo=FALSE, message=FALSE, warning=FALSE}
cat('###', "Immature Oligo CR", ' \n')
paged_table(immature_genelist_CR)
cat('\n \n')
cat('###', "Immature Oligo CONTROL", ' \n')
paged_table(immature_genelist_CONTROL)
cat('\n \n')
```

```{r,results='asis', echo=FALSE, message=FALSE, warning=FALSE}
cat('###', "Synaptic Oligo CR", ' \n')
paged_table(synaptic_genelist_CR)
cat('\n \n')
cat('###', "Synaptic Oligo CONTROL", ' \n')
paged_table(synaptic_genelist_CONTROL)
cat('\n \n')
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
david_CR <- DAVIDWebService$new(email='saumyapo@bu.edu',  url="https://david.ncifcrf.gov/webservice/services/DAVIDWebService.DAVIDWebServiceHttpSoap12Endpoint/")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#Get ID for each gene name (symbol) and run DAVID and store results for Mature Oligo CR condition
invisible(getIdTypes(david_CR))
CR_egids <- AnnotationDbi::select(org.Mmu.eg.db, genelist_CR$gene, "ENTREZID", "SYMBOL")
CR_result <- david_CR$addList(CR_egids$ENTREZID, idType = "ENTREZ_GENE_ID", listName = "Oligo CR", listType = "Gene")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
david_CONTROL <- DAVIDWebService$new(email='saumyapo@bu.edu',  url="https://david.ncifcrf.gov/webservice/services/DAVIDWebService.DAVIDWebServiceHttpSoap12Endpoint/")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#Get ID for each gene name (symbol) and run DAVID and store results for Mature Oligo CR condition
invisible(getIdTypes(david_CONTROL))
CONTROL_egids <- AnnotationDbi::select(org.Mmu.eg.db, genelist_CONTROL$gene, "ENTREZID", "SYMBOL")
CONTROL_result <- david_CONTROL$addList(CONTROL_egids$ENTREZID, idType = "ENTREZ_GENE_ID", listName = "Oligo CONTROL", listType = "Gene")
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
david_mature_CR <- DAVIDWebService$new(email='saumyapo@bu.edu',  url="https://david.ncifcrf.gov/webservice/services/DAVIDWebService.DAVIDWebServiceHttpSoap12Endpoint/")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#Get ID for each gene name (symbol) and run DAVID and store results for Mature Oligo CR condition
invisible(getIdTypes(david_mature_CR))
mature_CR_egids <- AnnotationDbi::select(org.Mmu.eg.db, mature_genelist_CR$gene, "ENTREZID", "SYMBOL")
mature_CR_result <- david_mature_CR$addList(mature_CR_egids$ENTREZID, idType = "ENTREZ_GENE_ID", listName = "Mature Oligo CR", listType = "Gene")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
david_mature_CONTROL <- DAVIDWebService$new(email='saumyapo@bu.edu',  url="https://david.ncifcrf.gov/webservice/services/DAVIDWebService.DAVIDWebServiceHttpSoap12Endpoint/")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#Get ID for each gene name (symbol) and run DAVID and store results for Mature Oligo CR condition
invisible(getIdTypes(david_mature_CONTROL))
mature_CONTROL_egids <- AnnotationDbi::select(org.Mmu.eg.db, mature_genelist_CONTROL$gene, "ENTREZID", "SYMBOL")
mature_CONTROL_result <- david_mature_CONTROL$addList(mature_CONTROL_egids$ENTREZID, idType = "ENTREZ_GENE_ID", listName = "Mature Oligo CONTROL", listType = "Gene")
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
david_immature_CR <- DAVIDWebService$new(email='saumyapo@bu.edu',  url="https://david.ncifcrf.gov/webservice/services/DAVIDWebService.DAVIDWebServiceHttpSoap12Endpoint/")
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
#Get ID for each gene name (symbol) and run DAVID and store results for Immature Oligo CR condition
invisible(getIdTypes(david_immature_CR))
immature_CR_egids <- AnnotationDbi::select(org.Mmu.eg.db, immature_genelist_CR$gene, "ENTREZID", "SYMBOL")
immature_CR_result <- david_immature_CR$addList(immature_CR_egids$ENTREZID, idType = "ENTREZ_GENE_ID", listName = "Immature Oligo CR", listType = "Gene")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
david_immature_CONTROL <- DAVIDWebService$new(email='saumyapo@bu.edu',  url="https://david.ncifcrf.gov/webservice/services/DAVIDWebService.DAVIDWebServiceHttpSoap12Endpoint/")
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
#Get ID for each gene name (symbol) and run DAVID and store results for Immature Oligo CONTROL condition
invisible(getIdTypes(david_immature_CONTROL))
immature_CONTROL_egids <- AnnotationDbi::select(org.Mmu.eg.db, immature_genelist_CONTROL$gene, "ENTREZID", "SYMBOL")
immature_CONTROL_result <- david_immature_CONTROL$addList(immature_CONTROL_egids$ENTREZID, idType = "ENTREZ_GENE_ID", listName = "Immature Oligo CONTROL", listType = "Gene")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
david_synaptic_CR <- DAVIDWebService$new(email='saumyapo@bu.edu',  url="https://david.ncifcrf.gov/webservice/services/DAVIDWebService.DAVIDWebServiceHttpSoap12Endpoint/")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#Get ID for each gene name (symbol) and run DAVID and store results for synaptic Oligo CR condition
invisible(getIdTypes(david_synaptic_CR))
synaptic_CR_egids <- AnnotationDbi::select(org.Mmu.eg.db, synaptic_genelist_CR$gene, "ENTREZID", "SYMBOL")
synaptic_CR_result <- david_synaptic_CR$addList(synaptic_CR_egids$ENTREZID, idType = "ENTREZ_GENE_ID", listName = "Synaptic Oligo CR", listType = "Gene")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
david_synaptic_CONTROL <- DAVIDWebService$new(email='saumyapo@bu.edu',  url="https://david.ncifcrf.gov/webservice/services/DAVIDWebService.DAVIDWebServiceHttpSoap12Endpoint/")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#Get ID for each gene name (symbol) and run DAVID and store results for synaptic Oligo CONTROL condition
invisible(getIdTypes(david_synaptic_CONTROL))
synaptic_CONTROL_egids <- AnnotationDbi::select(org.Mmu.eg.db, synaptic_genelist_CONTROL$gene, "ENTREZID", "SYMBOL")
synaptic_CONTROL_result <- david_synaptic_CONTROL$addList(synaptic_CONTROL_egids$ENTREZID, idType = "ENTREZ_GENE_ID", listName = "Synaptic Oligo CONTROL", listType = "Gene")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Convert the IDs back to gene names for easier understanding for Mature Oligo CR
setAnnotationCategories(david_CR, c("GOTERM_BP_DIRECT", "KEGG_PATHWAY", "REACTOME_PATHWAY"))
CR_gene_functions_df <- getFunctionalAnnotationChart(david_CR)
CR_gene_functions_df <- replacENTREZID(CR_gene_functions_df, CR_egids) 
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Convert the IDs back to gene names for easier understanding for Mature Oligo CONTROL
setAnnotationCategories(david_CONTROL, c("GOTERM_BP_DIRECT", "KEGG_PATHWAY", "REACTOME_PATHWAY"))
CONTROL_gene_functions_df <- getFunctionalAnnotationChart(david_CONTROL)
CONTROL_gene_functions_df <- replacENTREZID(CONTROL_gene_functions_df, CONTROL_egids) 
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Convert the IDs back to gene names for easier understanding for Mature Oligo CR
setAnnotationCategories(david_mature_CR, c("GOTERM_BP_DIRECT", "KEGG_PATHWAY", "REACTOME_PATHWAY"))
mature_CR_gene_functions_df <- getFunctionalAnnotationChart(david_mature_CR)
mature_CR_gene_functions_df <- replacENTREZID(mature_CR_gene_functions_df, mature_CR_egids) 
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Convert the IDs back to gene names for easier understanding for Mature Oligo CONTROL
setAnnotationCategories(david_mature_CONTROL, c("GOTERM_BP_DIRECT", "KEGG_PATHWAY", "REACTOME_PATHWAY"))
mature_CONTROL_gene_functions_df <- getFunctionalAnnotationChart(david_mature_CONTROL)
mature_CONTROL_gene_functions_df <- replacENTREZID(mature_CONTROL_gene_functions_df, mature_CONTROL_egids) 
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Convert the IDs back to gene names for easier understanding for Immature Oligo CR
setAnnotationCategories(david_immature_CR, c("GOTERM_BP_DIRECT", "KEGG_PATHWAY", "REACTOME_PATHWAY"))
immature_CR_gene_functions_df <- getFunctionalAnnotationChart(david_immature_CR)
immature_CR_gene_functions_df <- replacENTREZID(immature_CR_gene_functions_df, immature_CR_egids) 
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Convert the IDs back to gene names for easier understanding for Immature Oligo CONTROL
setAnnotationCategories(david_immature_CONTROL, c("GOTERM_BP_DIRECT", "KEGG_PATHWAY", "REACTOME_PATHWAY"))
immature_CONTROL_gene_functions_df <- getFunctionalAnnotationChart(david_immature_CONTROL)
immature_CONTROL_gene_functions_df <- replacENTREZID(immature_CONTROL_gene_functions_df, immature_CONTROL_egids) 
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Convert the IDs back to gene names for easier understanding for synaptic Oligo CR
setAnnotationCategories(david_synaptic_CR, c("GOTERM_BP_DIRECT", "KEGG_PATHWAY", "REACTOME_PATHWAY"))
synaptic_CR_gene_functions_df <- getFunctionalAnnotationChart(david_synaptic_CR)
synaptic_CR_gene_functions_df <- replacENTREZID(synaptic_CR_gene_functions_df, synaptic_CR_egids) 
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Convert the IDs back to gene names for easier understanding for synaptic Oligo CONTROL
setAnnotationCategories(david_synaptic_CONTROL, c("GOTERM_BP_DIRECT", "KEGG_PATHWAY", "REACTOME_PATHWAY"))
synaptic_CONTROL_gene_functions_df <- getFunctionalAnnotationChart(david_synaptic_CONTROL)
synaptic_CONTROL_gene_functions_df <- replacENTREZID(synaptic_CONTROL_gene_functions_df, synaptic_CONTROL_egids)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#Export Mature Oligo  results to xlsx
wb_results <- createWorkbook()
addWorksheet(wb_results, "Oligo CR")
writeData(wb_results, "Oligo CR", CR_gene_functions_df, rowNames = FALSE)
addWorksheet(wb_results, "Oligo CONTROL")
writeData(wb_results, "Oligo CONTROL", CONTROL_gene_functions_df, rowNames = FALSE)
saveWorkbook(wb_results, "oligo_DAVID_gene_functions.xlsx", overwrite = TRUE)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#Export Mature Oligo  results to xlsx
wb_mature_results <- createWorkbook()
addWorksheet(wb_mature_results, "Mature Oligo CR")
writeData(wb_mature_results, "Mature Oligo CR", mature_CR_gene_functions_df, rowNames = FALSE)
addWorksheet(wb_mature_results, "Mature Oligo CONTROL")
writeData(wb_mature_results, "Mature Oligo CONTROL", mature_CONTROL_gene_functions_df, rowNames = FALSE)
saveWorkbook(wb_mature_results, "mature_DAVID_gene_functions.xlsx", overwrite = TRUE)
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
#Export Immature Oligo  results to xlsx
wb_immature_results <- createWorkbook()
addWorksheet(wb_immature_results, "Immature Oligo CR")
writeData(wb_immature_results, "Immature Oligo CR", immature_CR_gene_functions_df, rowNames = FALSE)
addWorksheet(wb_immature_results, "Immature Oligo CONTROL")
writeData(wb_immature_results, "Immature Oligo CONTROL", immature_CONTROL_gene_functions_df, rowNames = FALSE)
saveWorkbook(wb_immature_results, "immature_DAVID_gene_functions.xlsx", overwrite = TRUE)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#Export Synaptic Oligo results to xlsx
wb_synaptic_results <- createWorkbook()
addWorksheet(wb_synaptic_results, "Synaptic Oligo CR")
writeData(wb_synaptic_results, "Synaptic Oligo CR", synaptic_CR_gene_functions_df, rowNames = FALSE)
addWorksheet(wb_synaptic_results, "Synaptic Oligo CONTROL")
writeData(wb_synaptic_results, "Synaptic Oligo CONTROL", synaptic_CONTROL_gene_functions_df, rowNames = FALSE)
saveWorkbook(wb_synaptic_results, "synaptic_DAVID_gene_functions.xlsx", overwrite = TRUE)
```


## `r colorize('Results for DAVID Enrichment Analysis', 'blue')` {.tabset}

### Understanding output of analysis

The output of the enrichment analysis is an excel file which has 13 headers and explanation on each of the headers can be found below:

1. Category: The type of functional annotation. Example, GOTERM_BP_DIRECT, indicating direct Gene Ontology terms related to biological processes.

2. Term: The specific GO term and its definition. For example, GO:0045944~positive regulation of transcription from RNA polymerase II promoter.

3. Count: The number of genes from the input list associated with the given GO term. A higher count indicates more genes in your input list are associated with the GO term, suggesting stronger enrichment.

4. X.: The combined score or enrichment score, often a measure of the statistical significance and relevance of the GO term. Higher enrichment scores indicate stronger relevance of the GO term to the input gene list. Generally, an enrichment score > 1.3 is considered significant.

5. PValue: The p-value indicating the statistical significance of the enrichment of the GO term. Lower p-values suggest higher statistical significance. 

6. Genes: The list of genes from the input set that are associated with the GO term. For instance, genes like Il10, Kdm6b, and Junb are associated with GO:0045944.

7. List.Total: The total number of genes in the input list used for the analysis.

8. Pop.Hits: The number of genes in the background population that are associated with the GO term. 

9. Pop.Total: The total number of genes in the background population considered in the analysis. 

10. Fold.Enrichment: The ratio of the observed frequency of the GO term in the input list to the expected frequency in the background population. It indicates how much more common the term is in the input list compared to the background. For example, 2.683442029 means the term is approximately 2.68 times more common in the input list than expected by chance. Higher fold enrichment values indicate that the GO term is more over-represented in the input list compared to the background. Generally, but a fold enrichment > 2 is typically considered noteworthy.

11. Bonferroni: The p-value adjusted for multiple testing using the Bonferroni correction. It controls the family-wise error rate. The lower the number the better.

12. Benjamini: The p-value adjusted for multiple testing using the Benjamini-Hochberg procedure, which controls the false discovery rate (FDR). The lower the number the better.

13. FDR: The false discovery rate, another method to correct for multiple testing, showing the expected proportion of false positives among the significant results. The lower the number the better.


```{r,results='asis', echo=FALSE, message=FALSE, warning=FALSE}
cat('###', "Oligo CR", ' \n')
paged_table(CR_gene_functions_df)
cat(' \n \n')
cat('###', "Oligo CONTROL", ' \n')
paged_table(CONTROL_gene_functions_df)
cat(' \n \n')
cat('###', "Mature Oligo CR", ' \n')
paged_table(mature_CR_gene_functions_df)
cat(' \n \n')
cat('###', "Mature Oligo CONTROL", ' \n')
paged_table(mature_CONTROL_gene_functions_df)
cat(' \n \n')
cat('###', "Immature Oligo CR", ' \n')
paged_table(immature_CR_gene_functions_df)
cat(' \n \n')
cat('###', "Immature Oligo CONTROL", ' \n')
paged_table(immature_CONTROL_gene_functions_df)
cat(' \n \n')
cat('###', "Synaptic Oligo CR", ' \n')
paged_table(synaptic_CR_gene_functions_df)
cat(' \n \n')
cat('###', "Synaptic Oligo CONTROL", ' \n')
paged_table(synaptic_CONTROL_gene_functions_df)
cat(' \n \n')
```



```{r, echo=FALSE, message=FALSE, warning=FALSE}
CR_plot = plotRrvgo(CR_gene_functions_df)
CONTROL_plot = plotRrvgo(CONTROL_gene_functions_df)
mature_CR_plot = plotRrvgo(mature_CR_gene_functions_df)
immature_CR_plot = plotRrvgo(immature_CR_gene_functions_df)
synaptic_CR_plot = plotRrvgo(synaptic_CR_gene_functions_df)
mature_CONTROL_plot = plotRrvgo(mature_CONTROL_gene_functions_df)
immature_CONTROL_plot = plotRrvgo(immature_CONTROL_gene_functions_df)
synaptic_CONTROL_plot = plotRrvgo(synaptic_CONTROL_gene_functions_df)
```

## `r colorize('Plotting Gene Enrichment results using treemap and scatterplot', 'blue')` {.tabset}

### Oligo CR {.tabset}
```{r,results='asis', echo=FALSE, message=FALSE, warning=FALSE}
cat('####', "treemapPlot", ' \n')
treemapPlot(CR_plot$second)
cat(' \n \n')
cat('####', "scatterPlot", ' \n')
print(scatterPlot(CR_plot$first, CR_plot$second))
cat(' \n \n')
```

### Oligo CONTROL {.tabset}
```{r,results='asis', echo=FALSE, message=FALSE, warning=FALSE}
cat('####', "treemapPlot", ' \n')
treemapPlot(CONTROL_plot$second)
cat(' \n \n')
cat('####', "scatterPlot", ' \n')
print(scatterPlot(CONTROL_plot$first, CONTROL_plot$second))
cat(' \n \n')
```

### Mature Oligo CR {.tabset}
```{r,results='asis', echo=FALSE, message=FALSE, warning=FALSE}
cat('####', "treemapPlot", ' \n')
treemapPlot(mature_CR_plot$second)
cat(' \n \n')
cat('####', "scatterPlot", ' \n')
print(scatterPlot(mature_CR_plot$first, mature_CR_plot$second))
cat(' \n \n')
```  

### Mature Oligo CONTROL {.tabset}
```{r,results='asis', echo=FALSE, message=FALSE, warning=FALSE}
cat('####', "treemapPlot", ' \n')
treemapPlot(mature_CONTROL_plot$second)
cat(' \n \n')
cat('####', "scatterPlot", ' \n')
print(scatterPlot(mature_CONTROL_plot$first, mature_CONTROL_plot$second))
cat(' \n \n')
``` 

### Immature Oligo CR {.tabset}
```{r,results='asis', echo=FALSE, message=FALSE, warning=FALSE}
cat('####', "treemapPlot", ' \n')
treemapPlot(immature_CR_plot$second)
cat(' \n \n')
cat('####', "scatterPlot", ' \n')
print(scatterPlot(immature_CR_plot$first, immature_CR_plot$second))
cat(' \n \n')  
```

### Immature Oligo CONTROL {.tabset}
```{r,results='asis', echo=FALSE, message=FALSE, warning=FALSE}
cat('####', "treemapPlot", ' \n')
treemapPlot(immature_CONTROL_plot$second)
cat(' \n \n')
cat('####', "scatterPlot", ' \n')
print(scatterPlot(immature_CONTROL_plot$first, immature_CONTROL_plot$second))
cat(' \n \n')  
```

### Synaptic Oligo CR {.tabset}
```{r,results='asis', echo=FALSE, message=FALSE, warning=FALSE}
cat('####', "treemapPlot", ' \n')
treemapPlot(synaptic_CR_plot$second)
cat(' \n \n')
cat('####', "scatterPlot", ' \n')
print(scatterPlot(synaptic_CR_plot$first, synaptic_CR_plot$second))
cat(' \n \n')
```

### Synaptic Oligo CONTROL {.tabset}
```{r,results='asis', echo=FALSE, message=FALSE, warning=FALSE}
cat('####', "treemapPlot", ' \n')
treemapPlot(synaptic_CONTROL_plot$second)
cat(' \n \n')
cat('####', "scatterPlot", ' \n')
print(scatterPlot(synaptic_CONTROL_plot$first, synaptic_CONTROL_plot$second))
cat(' \n \n')
```
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Function to remove GO IDs from the Term column
remove_go_id <- function(term) {
  str_replace(term, "^GO:[0-9]+~", "")
}
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Read Entire Oligo CR DAVID results
CR_david <- read_excel("oligo_DAVID_gene_functions.xlsx", sheet="Oligo CR")

# Remove GO IDs from the Term column
CR_david$Term <- sapply(CR_david$Term, remove_go_id)

# Filter for top 10 pathways based on P-value
top_CR_results <- CR_david %>%
  arrange(PValue) %>%
  head(10)

# Create a horizontal bar plot of top 10 enriched pathways
plot_top_CR_pathways <- ggplot(top_CR_results, aes(x = reorder(Term, -PValue), y = -log10(PValue))) +
  geom_bar(stat = "identity", fill = "skyblue") +
  coord_flip() +
  labs(title = "Top 10 Pathways Oligo CR DAVID", x = "Pathway", y = "-log10(P-value)") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 6))
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Read Mature Oligo CONTROL DAVID results
CONTROL_david <- read_excel("oligo_DAVID_gene_functions.xlsx", sheet="Oligo CONTROL")

# Remove GO IDs from the Term column
CONTROL_david$Term <- sapply(CONTROL_david$Term, remove_go_id)

# Filter for top 10 pathways based on P-value
top_CONTROL_results <- CONTROL_david %>%
  arrange(PValue) %>%
  head(10)

# Create a horizontal bar plot of top 10 enriched pathways
plot_top_CONTROL_pathways <- ggplot(top_CONTROL_results, aes(x = reorder(Term, -PValue), y = -log10(PValue))) +
  geom_bar(stat = "identity", fill = "darkred") +
  coord_flip() +
  labs(title = "Top 10 Pathways Oligo CONTROL DAVID", x = "Pathway", y = "-log10(P-value)") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 6))
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Read Mature Oligo CR DAVID results
mature_CR_david <- read_excel("mature_DAVID_gene_functions.xlsx", sheet="Mature Oligo CR")

# Remove GO IDs from the Term column
mature_CR_david$Term <- sapply(mature_CR_david$Term, remove_go_id)

# Filter for top 10 pathways based on P-value
top_mature_CR_results <- mature_CR_david %>%
  arrange(PValue) %>%
  head(10)

# Create a horizontal bar plot of top 10 enriched pathways
plot_top_mature_CR_pathways <- ggplot(top_mature_CR_results, aes(x = reorder(Term, -PValue), y = -log10(PValue))) +
  geom_bar(stat = "identity", fill = "skyblue") +
  coord_flip() +
  labs(title = "Top 10 Pathways Mature Oligo CR DAVID", x = "Pathway", y = "-log10(P-value)") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 6))
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Read Mature Oligo CONTROL DAVID results
mature_CONTROL_david <- read_excel("mature_DAVID_gene_functions.xlsx", sheet="Mature Oligo CONTROL")

# Remove GO IDs from the Term column
mature_CONTROL_david$Term <- sapply(mature_CONTROL_david$Term, remove_go_id)

# Filter for top 10 pathways based on P-value
top_mature_CONTROL_results <- mature_CONTROL_david %>%
  arrange(PValue) %>%
  head(10)

# Create a horizontal bar plot of top 10 enriched pathways
plot_top_mature_CONTROL_pathways <- ggplot(top_mature_CONTROL_results, aes(x = reorder(Term, -PValue), y = -log10(PValue))) +
  geom_bar(stat = "identity", fill = "darkred") +
  coord_flip() +
  labs(title = "Top 10 Pathways Mature Oligo CONTROL DAVID", x = "Pathway", y = "-log10(P-value)") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 6))
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Read Immature Oligo CR DAVID results
immature_CR_david <- read_excel("immature_DAVID_gene_functions.xlsx", sheet="Immature Oligo CR")

# Remove GO IDs from the Term column
immature_CR_david$Term <- sapply(immature_CR_david$Term, remove_go_id)

# Filter for top 10 pathways based on P-value
top_immature_CR_results <- immature_CR_david %>%
  arrange(PValue) %>%
  head(10)

# Create a horizontal bar plot of top 10 enriched pathways
plot_top_immature_CR_pathways <- ggplot(top_immature_CR_results, aes(x = reorder(Term, -PValue), y = -log10(PValue))) +
  geom_bar(stat = "identity", fill = "skyblue") +
  coord_flip() +
  labs(title = "Top 10 Pathways Immature Oligo CR DAVID", x = "Pathway", y = "-log10(P-value)") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 6))
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Read Immature Oligo CONTROL DAVID results
immature_CONTROL_david <- read_excel("immature_DAVID_gene_functions.xlsx", sheet="Immature Oligo CONTROL")

# Remove GO IDs from the Term column
immature_CONTROL_david$Term <- sapply(immature_CONTROL_david$Term, remove_go_id)

# Filter for top 10 pathways based on P-value
top_immature_CONTROL_results <- immature_CONTROL_david %>%
  arrange(PValue) %>%
  head(10)

# Create a horizontal bar plot of top 10 enriched pathways
plot_top_immature_CONTROL_pathways <- ggplot(top_immature_CONTROL_results, aes(x = reorder(Term, -PValue), y = -log10(PValue))) +
  geom_bar(stat = "identity", fill = "darkred") +
  coord_flip() +
  labs(title = "Top 10 Pathways Immature Oligo CONTROL DAVID", x = "Pathway", y = "-log10(P-value)") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 6))
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Read Synaptic Oligo CR DAVID results
synaptic_CR_david <- read_excel("synaptic_DAVID_gene_functions.xlsx", sheet="Synaptic Oligo CR")

# Remove GO IDs from the Term column
synaptic_CR_david$Term <- sapply(synaptic_CR_david$Term, remove_go_id)

# Filter for top 10 pathways based on P-value
top_synaptic_CR_results <- synaptic_CR_david %>%
  arrange(PValue) %>%
  head(10)

# Create a horizontal bar plot of top 10 enriched pathways
plot_top_synaptic_CR_pathways <- ggplot(top_synaptic_CR_results, aes(x = reorder(Term, -PValue), y = -log10(PValue))) +
  geom_bar(stat = "identity", fill = "skyblue") +
  coord_flip() +
  labs(title = "Top 10 Pathways Synaptic Oligo CR DAVID", x = "Pathway", y = "-log10(P-value)") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 6))
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Read Synaptic Oligo CONTROL DAVID results
synaptic_CONTROL_david <- read_excel("synaptic_DAVID_gene_functions.xlsx", sheet="Synaptic Oligo CONTROL")

# Remove GO IDs from the Term column
synaptic_CONTROL_david$Term <- sapply(synaptic_CONTROL_david$Term, remove_go_id)

# Filter for top 10 pathways based on P-value
top_synaptic_CONTROL_results <- synaptic_CONTROL_david %>%
  arrange(PValue) %>%
  head(10)

# Create a horizontal bar plot of top 10 enriched pathways
plot_top_synaptic_CONTROL_pathways <- ggplot(top_synaptic_CONTROL_results, aes(x = reorder(Term, -PValue), y = -log10(PValue))) +
  geom_bar(stat = "identity", fill = "darkred") +
  coord_flip() +
  labs(title = "Top 10 Pathways Synaptic Oligo CONTROL DAVID", x = "Pathway", y = "-log10(P-value)") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 6))
```

## `r colorize('Pathway Plot of Top 10 pathways based on p-value', 'blue')` {.tabset}
The pathways upregulated in CR are downregulated in CONTROL and vice versa. So the top pathways plotted for each group (CR and CONTROL) are the upregulated pathways for that specific group while also being the downregulated pathways for the other group.

```{r,results='asis', echo=FALSE, message=FALSE, warning=FALSE,fig.width=10,fig.height=6}
# Display DAVID barplots
cat('###', "Oligo CR", ' \n')
print(plot_top_CR_pathways)
cat(' \n \n')
cat('###', "Oligo CONTROL", ' \n')
print(plot_top_CONTROL_pathways)
cat(' \n \n')
cat('###', "Mature Oligo CR", ' \n')
print(plot_top_mature_CR_pathways)
cat(' \n \n')
cat('###', "Mature Oligo CONTROL", ' \n')
print(plot_top_mature_CONTROL_pathways)
cat(' \n \n')
cat('###', "Immature Oligo CR", ' \n')
print(plot_top_immature_CR_pathways)
cat(' \n \n')
cat('###', "Immature Oligo CONTROL", ' \n')
print(plot_top_immature_CONTROL_pathways)
cat(' \n \n')
cat('###', "Synaptic Oligo CR", ' \n')
print(plot_top_synaptic_CR_pathways)
cat(' \n \n')
cat('###', "Synaptic Oligo CONTROL", ' \n')
print(plot_top_synaptic_CONTROL_pathways)
cat(' \n \n')
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
pdf(paste("pdf_oligo_DAVID_pathway_plots.pdf"), width=10, height=8, useDingbats = FALSE)
print(plot_top_CR_pathways)
print(plot_top_CONTROL_pathways)
dev.off()
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
pdf(paste("pdf_mature_oligo_DAVID_pathway_plots.pdf"), width=10, height=8, useDingbats = FALSE)
print(plot_top_mature_CR_pathways)
print(plot_top_mature_CONTROL_pathways)
dev.off()
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
pdf(paste("pdf_immature_oligo_DAVID_pathway_plots.pdf"), width=10, height=8, useDingbats = FALSE)
print(plot_top_immature_CR_pathways)
print(plot_top_immature_CONTROL_pathways)
dev.off()
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
pdf(paste("pdf_synaptic_oligo_DAVID_pathway_plots.pdf"), width=10, height=8, useDingbats = FALSE)
print(plot_top_synaptic_CR_pathways)
print(plot_top_synaptic_CONTROL_pathways)
dev.off()
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
pdf(paste("pdf_oligo_3_subclusters_CR_DAVID_pathway_plots.pdf"), width=10, height=8, useDingbats = FALSE)
print(plot_top_immature_CR_pathways)
print(plot_top_mature_CR_pathways)
print(plot_top_synaptic_CR_pathways)
dev.off()
```