---
title: "T Cell Clustering"
author: "Saumya"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# Markdown parameters
set_echo = FALSE
set_eval = FALSE
```

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# Function to set markdown text color
colorize <- function(x, color) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{%s}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'>%s</span>", color,
      x)
  } else x
}
```


```{r, include=FALSE}
# Set height of chunk output
options(width = 60)
local({
  hook_output <- knitr::knit_hooks$get('output')
  knitr::knit_hooks$set(output = function(x, options) {
    if (!is.null(options$max.height)) options$attr.output <- c(
      options$attr.output,
      sprintf('style="max-height: %s;"', options$max.height)
    )
    hook_output(x, options)
  })
})
```

```{r, results = FALSE, echo=set_echo, message=FALSE, warning=FALSE}
# Load packages
library(dplyr)
library(Seurat)
library(patchwork)
library(ggplot2)
library(cowplot)
library(rmarkdown)
library(SummarizedExperiment)
library(htmlwidgets)
library(htmltools)
library(shiny)
library(openxlsx)
library(scDblFinder)
library(hdf5r)
library(harmony, lib.loc ="/hpc/users/pothus02/.Rlib")
library(DT)
```

<!-- Information of Used Library -->
### `r colorize('Information of Used R Libraries', 'blue')` 

The following R libraries were used to process the data and produce this report.

R: `r R.Version()$version.string`,

dplyr: `r packageVersion('dplyr')`, 

Seurat: `r packageVersion('Seurat')`, 

patchwork: `r packageVersion('patchwork')`, 

ggplot2: `r packageVersion('ggplot2')`, 

cowplot: `r packageVersion('cowplot')`,

scDblFinder: `r packageVersion('scDblFinder')`

```{r, results = FALSE, echo=set_echo, message=FALSE, warning=FALSE}
suffix <- "Seurat"
sampleNames <- c("DART002LC_BL", "DART003LC", "DART005LC", "DART006LC", "DART007LC", "HIV018_CA", "HIV022_CA", "HIV032_CA", "HIV058_LP", "HIV063_LP", "HIV075_LP", "HIV076_LP", "HIV077_LP", "HIV078_LP", "HIV080_LP", "HIV081_LP", "HIV082_LP", "HIV083_LC", "HIV084_LP", "HIV086_LC", "HIV087_LP", "HIV090_LC", "HIV091_LC", "HIV092_LC", "HIV093_LC", "HIV094_LC", "HIV098_LC", "NV144_LP", "NV157_LP_LC", "NV159_LP_LC", "NV161_LC", "NV162_LP_LC", "NV186_LP", "NV201_LP", "NV203_LP", "NV205_LP", "NV212_LC", "NV213_LC", "NV216_LP", "NV222_LP", "NV225_LP", "NV230_LP", "NV234_LC", "NV244_LC", "NV245_LC")

SAMPLE_GROUP <- c("HIV", "HIV", "HIV","HIV","HIV","HIV","HIV","HIV","HIV","HIV","HIV","HIV","HIV","HIV","HIV","HIV","HIV","HIV","HIV","HIV","HIV","HIV","HIV","HIV","HIV","HIV","HIV","Control","Control","Control","Control","Control","Control","Control","Control","Control","Control","Control", "Control", "Control", "Control","Control","Control","Control","Control")
data_dir <- ("../raw_h5/")
projectName <- "Immune HIV vs Control"
h5_string <- "_filtered_feature_bc_matrix.h5"

#Seurat parameters
variables.features <- 2000
nPCs <- 50
```

<!-- # `r colorize('T Cell Subcluster ', 'blue')` -->

```{r, results = FALSE, echo=set_echo, message=FALSE, warning=FALSE}
# data.combined <- readRDS("../RDS/data_44_umap_33_clusters.RDS")
# tcell_subset <- subset(data.combined, idents = c("2","3","6","10","20","29"))
```

```{r, eval=FALSE, results = FALSE, echo=set_echo, message=FALSE, warning=FALSE}
# tcell_subset <- NormalizeData(tcell_subset, normalization.method = "LogNormalize", scale.factor = 10000)
# tcell_subset <- FindVariableFeatures(tcell_subset, selection.method = "vst", nfeatures = 2000)
# tcell_subset <- ScaleData(tcell_subset, verbose = FALSE)
# 
# #dimension reduction
# tcell_subset <- RunPCA(tcell_subset, npcs = 50, verbose = FALSE)
# # Run Harmony for batch correction
# tcell_subset <- RunHarmony(tcell_subset, "sample_names")
# tcell_subset <- RunUMAP(tcell_subset, reduction = "harmony", dims = 1:50)
```

```{r, eval = FALSE, results = FALSE, echo=FALSE, message=FALSE, warning=FALSE}
#saveRDS(tcell_subset, file = paste("../RDS/tcell_subset_umap.RDS"))
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# tcell_subset <- readRDS("../RDS/tcell_subset_umap.RDS")
# 
# cluster_annotations <- c("2" = "T cells",
#                          "3" = "T cells",
#                          "6" = "T cells",
#                          "10" = "T cells",
#                          "20" = "T cells",
#                          "29" = "T cells")
# 
# #Apply the new annotations to your Seurat object
# tcell_subset$cluster_label <- plyr::mapvalues(
#   Idents(tcell_subset),
#    from = names(cluster_annotations),
#   to = cluster_annotations
# )
# 
# #Update the identities in the Seurat object
# Idents(tcell_subset) <- tcell_subset$cluster_label
```


# `r colorize('T cells before clustering', 'blue')`

```{r,results='asis', echo=FALSE, message=FALSE, warning=FALSE}
tcell_subset <- readRDS("../RDS/tcell_subset_umap.RDS")
Idents(tcell_subset) <- "cluster_label"
DimPlot(tcell_subset, reduction = "umap") + ggtitle ("T cells Before Clustering") 
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Extract raw counts for the neutrophil cluster (cluster 5)
data.combined <- readRDS("../RDS/data_44_umap_33_clusters.RDS")
tcell_subset <- readRDS("../RDS/tcell_subset_umap.RDS")

# Get the raw count of neutrophil cells
T_count <- length(Cells(tcell_subset))

# Get the total number of cells in the dataset
total_cells <- length(Cells(data.combined))

# Calculate the percentage of neutrophils in the dataset
T_percentage <- (T_count / total_cells) * 100

# Create a summary table
summary_table <- data.frame(
  "Cell Type" = c("T Cells", "Total Cells"),
  "Raw Count" = c(T_count, total_cells),
  "Percentage" = c(T_percentage, 100)
)
```

# Cell Count for T Cells {.tabset}
```{r, results='asis', echo=FALSE, message=FALSE, warning=FALSE}
# Display the summary table
paged_table(summary_table)
```


# `r colorize('Sub-clustering for T cells', 'blue')`

```{r, eval = FALSE, echo=FALSE, message = FALSE, results= FALSE, warning=FALSE}
# Do clustering
# tcell_subset <- readRDS("../RDS/tcell_subset_umap.RDS")
# tcell_subset <- FindNeighbors(tcell_subset, reduction = "harmony", dims = 1:50)
# tcell_subset <- FindClusters(tcell_subset)

#ncol(tcell_subset)
```

```{r,results='asis', echo=FALSE, message=FALSE, warning=FALSE}

tcell_subset <- readRDS("../RDS/data_tcell_umap_19_clusters.RDS")
Idents(tcell_subset) <- "seurat_clusters"
# Check clustering UMAP
DimPlot(tcell_subset, reduction = "umap", label = TRUE, repel = TRUE, raster=FALSE) + ggtitle("Seurat Clusters")

#ElbowPlot(tcell_subset, ndims = 50)
```

```{r, eval = FALSE, echo=FALSE, message = FALSE, results= FALSE, warning=FALSE}
# Save clustering data
#saveRDS(tcell_subset, file = paste("../RDS/data_tcell_umap_19_clusters.RDS") )
```

# `r colorize('DEGS for 19 T Cell sub-clusters', 'blue')`

```{r, eval = FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# Find All Markers
# tcell_subset <- readRDS("../RDS/data_tcell_umap_19_clusters.RDS")
# DefaultAssay(tcell_subset) <- "RNA"
# Idents(tcell_subset) <- "condition"
# tcell_subset <- JoinLayers(tcell_subset)
# data.markers <- FindAllMarkers(tcell_subset, only.pos = FALSE, min.pct = 0.25, logfc.threshold = 0.25)
# # Save markers
# write.csv(data.markers, file='../csvs/tcells/condition/data_tcell_all_markers_condition.csv')
```

```{r, eval = FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# Find All Markers
# tcell_subset <- readRDS("../RDS/data_tcell_umap_19_clusters.RDS")
# DefaultAssay(tcell_subset) <- "RNA"
# Idents(tcell_subset) <- "seurat_clusters"
# tcell_subset <- JoinLayers(tcell_subset)
# data.markers <- FindAllMarkers(tcell_subset, only.pos = FALSE, min.pct = 0.25, logfc.threshold = 0.25)
# # Save markers
# write.csv(data.markers, file='../csvs/data_tcell_all_markers_19_clusters.csv')
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Read Markers .csv
data.markers = read.csv('../csvs/tcells/data_tcell_all_markers_19_clusters.csv')
data.markers$pct.diff = data.markers$pct.1 - data.markers$pct.2
data.markers$pct.ratio = data.markers$pct.1 / data.markers$pct.2
avglog_p = data.markers %>% group_by(cluster) %>% slice_max(n = 20, order_by = avg_log2FC)
avglog_n = data.markers %>% group_by(cluster) %>% slice_min(n = 20, order_by = avg_log2FC)
pctdiff_p = data.markers %>% group_by(cluster) %>% slice_max(n = 20, order_by = pct.diff)
pctdiff_n = data.markers %>% group_by(cluster) %>% slice_min(n = 20, order_by = pct.diff)
pctratio_p = data.markers %>% group_by(cluster) %>% slice_max(n = 20, order_by = pct.ratio)
pctratio_n = data.markers %>% group_by(cluster) %>% slice_min(n = 20, order_by = pct.ratio)
```


```{r, eval = FALSE, echo=FALSE, message=FALSE, warning=FALSE}

positive.markers = read.csv('../csvs/tcells/data_tcell_all_markers_19_clusters.csv')

positive.markers$pct.diff = positive.markers$pct.1 - positive.markers$pct.2
positive.markers$pct.ratio = positive.markers$pct.1 / positive.markers$pct.2

# Keep positive markers only
positive.markers <- positive.markers[positive.markers$pct.diff > 0.0,]

positive.markers = positive.markers %>% group_by(cluster) %>% slice_max(n = 200, order_by = pct.diff)


write.csv(positive.markers, file='../csvs/tcells/data_tcell_pos_markers_19_clusters.csv')

```


```{r, eval = FALSE, echo=FALSE, message=FALSE, warning=FALSE}
paged_table(positive.markers)
```

## Positive markers {.tabset}
While downloading tables please make sure to change filename, otherwise it might overwrite if downloading multiple tables. Make sure to include ".csv" after changing filename to retain format.

### Top 20 Markers (avg_log2FC)
```{r,results='asis', echo=FALSE, message=FALSE, warning=FALSE}
# Create a table with a CSV download option
datatable(
  avglog_p, 
  extensions = 'Buttons', 
  options = list(
    dom = 'Bfrtip',
    buttons = list('csv'),
    filter = "top"  # Ensures a search box appears
  )
)
```

### Top 20 Markers (pctdiff)
```{r,results='asis', echo=FALSE, message=FALSE, warning=FALSE}
datatable(
  pctdiff_p, 
  extensions = 'Buttons', 
  options = list(
    dom = 'Bfrtip',
    buttons = list('csv'),
    filter = "top"
  )
)
```

### Top 20 Markers (pct.ratio)
```{r,results='asis', echo=FALSE, message=FALSE, warning=FALSE}
datatable(
  pctratio_p, 
  extensions = 'Buttons', 
  options = list(
    dom = 'Bfrtip',
    buttons = list('csv', 'search')
  )
)
```

## Negative markers {.tabset}
While downloading tables please make sure to change filename, otherwise it might overwrite if downloading multiple tables. Make sure to include ".csv" after changing filename to retain format.

### Top 20 Markers (avg_log2FC)
```{r,results='asis', echo=FALSE, message=FALSE, warning=FALSE}
# Create a table with a CSV download option
datatable(
  avglog_n, 
  extensions = 'Buttons', 
  options = list(
    dom = 'Bfrtip',
    buttons = list('csv'),
    filter = "top"  # Ensures a search box appears
  )
)
```

### Top 20 Markers (pctdiff)
```{r,results='asis', echo=FALSE, message=FALSE, warning=FALSE}
datatable(
  pctdiff_n, 
  extensions = 'Buttons', 
  options = list(
    dom = 'Bfrtip',
    buttons = list('csv'),
    filter = "top"
  )
)
```

### Top 20 Markers (pct.ratio)
```{r,results='asis', echo=FALSE, message=FALSE, warning=FALSE}
datatable(
  pctratio_n, 
  extensions = 'Buttons', 
  options = list(
    dom = 'Bfrtip',
    buttons = list('csv', 'search')
  )
)
```


## All markers
While downloading tables please make sure to change filename, otherwise it might overwrite if downloading multiple tables. Make sure to include ".csv" after changing filename to retain format.

```{r,results='asis', echo=FALSE, message=FALSE, warning=FALSE}
# Create a table with a CSV download option
datatable(
  data.markers, 
  extensions = 'Buttons', 
  options = list(
    dom = 'Bfrtip',
    buttons = list('csv')
  )
)
```




```{r, results = FALSE, eval = FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# tcell_subset <- readRDS("../RDS/data_tcell_umap_19_clusters.RDS")
# 
# cluster_annotations <- c("0" = "Naïve CD4+",
#                          "1" = "Cytotoxic CD8+",
#                          "2" = "Effector CD8+",
#                          "3" = "Th17-like CD4+",
#                          "4" = "Tregs",
#                          "5" = "TFH",
#                          "6" = "Tcm",
#                          "7" = "Tcm",
#                          "8" = "Tem",
#                          "9" = "Tem",
#                          "10" = "TRM",
#                          "11" = "Exhausted CD8+",
#                          "12" = "GDT",
#                          "13" = "NK T cells",
#                          "14" = "Proliferating",
#                          "15" = "Proliferating",
#                          "16" = "Proliferating",
#                          "17" = "MAIT",
#                          "18" = "Innate-like CD8+")
# 
# 
# 
# #Apply the new annotations to your Seurat object
# tcell_subset$cluster_label <- plyr::mapvalues(
#   Idents(tcell_subset),
#    from = names(cluster_annotations),
#   to = cluster_annotations
# )
# 
# #Update the identities in the Seurat object
# Idents(tcell_subset) <- tcell_subset$cluster_label
# 
# saveRDS(tcell_subset, "../RDS/data_tcell_umap_19_clusters.RDS")
```


# `r colorize('Annotated T cell clusters', 'blue')`

```{r,results='asis', echo=FALSE, message=FALSE, warning=FALSE}
tcell_subset <- readRDS("../RDS/data_tcell_umap_19_clusters.RDS")
Idents(tcell_subset) <- "cluster_label"
# Check clustering UMAP
DimPlot(tcell_subset, reduction = "umap", label = TRUE, repel = TRUE, raster=FALSE) + ggtitle("Seurat Clusters")

#ElbowPlot(tcell_subset, ndims = 50)
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Extract raw counts for the neutrophil cluster (cluster 5)
#tcell_subset <- readRDS("../RDS/data_tcell_umap_19_clusters.RDS")

treg_cells <- WhichCells(tcell_subset, ident = "Tregs")

# Get the raw count of neutrophil cells
treg_count <- length(treg_cells)

# Get the total number of cells in the dataset
total_cells <- length(Cells(tcell_subset))

# Calculate the percentage of neutrophils in the dataset
treg_percentage <- (treg_count / total_cells) * 100

# Create a summary table
summary_table <- data.frame(
  "Cell Type" = c("Tregs", "Total Cells"),
  "Raw Count" = c(treg_count, total_cells),
  "Percentage" = c(treg_percentage, 100)
)
```

# Cell Count for Tregs {.tabset}
```{r, results='asis', echo=FALSE, message=FALSE, warning=FALSE}
# Display the summary table
paged_table(summary_table)
```


```{r, results = FALSE, eval = FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# # Read Seurat object
# tcell_subset <- readRDS("../RDS/data_tcell_umap_19_clusters.RDS")
# 
# # Ensure RNA assay is set
# DefaultAssay(tcell_subset) <- "RNA"
# Idents(tcell_subset) <- "seurat_clusters"
# 
# tcell_subset <- JoinLayers(tcell_subset)
# 
# # Store subcluster identities (backup)
# tcell_subset$subcluster <- Idents(tcell_subset)
# 
# # Set condition as active identity
# Idents(tcell_subset) <- "condition"
# 
# # Create an empty list to store results
# markers_list <- list()
# 
# # Loop over each subcluster
# for (cluster in unique(tcell_subset$subcluster)) {
#   # Subset data for the current subcluster
#   subcluster_data <- subset(tcell_subset, subset = subcluster == cluster)
# 
#   # Set condition as active identity
#   Idents(subcluster_data) <- "condition"
# 
#   # Find all markers between HIV and Control within this subcluster
#   markers <- FindAllMarkers(subcluster_data, only.pos = FALSE, min.pct = 0.1, logfc.threshold = 0)
#   markers$p_val_adj <- p.adjust(markers$p_val, method = "fdr")
# 
#   # Store results with subcluster name
#   markers_list[[paste0("Tcell_subcluster_", cluster)]] <- markers
# }
# 
# # Save results to CSV
# for (name in names(markers_list)) {
#   write.csv(markers_list[[name]], file=paste0("../csvs/tcells/condition/", name, "_HIV_vs_Control.csv"))
# }
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Get list of all subcluster DEG files
subcluster_files <- list.files("../csvs/tcells/condition/", pattern = "^Tcell_", full.names = TRUE)

# Initialize empty data frame to store counts
deg_summary <- data.frame(Subcluster = character(), Num_Sig_DEGs = numeric(), stringsAsFactors = FALSE)

# Loop through each file and count significant DEGs
for (file in subcluster_files) {
  # Read data
  deg_data <- read.csv(file)
  
  # Filter for significant DEGs
  sig_degs <- deg_data %>% filter(p_val_adj < 0.05 & abs(avg_log2FC) > 1) %>%
  filter(!grepl("^MT", gene) & !grepl("^RB[SL]", gene))  # Exclude mitochondrial and ribosomal genes
  
  # Extract subcluster name from file name
  subcluster_name <- gsub("../csvs/tcells/condition/|_HIV_vs_Control.csv", "", file)
  
  # Store result
  deg_summary <- rbind(deg_summary, data.frame(Subcluster = subcluster_name, Num_Sig_DEGs = nrow(sig_degs)))
}

# Save summary to CSV
#write.csv(deg_summary, "../csvs/DEG_Summary.csv", row.names = FALSE)
```

# `r colorize('Number of significant DEGs (HIV vs Control) in each subcluster', 'blue')` {.tabset}

- p_val_adj < 0.05, and abs(log2fc) > 1

- Not counting mitochondrial and ribosomal genes. They are still present in the data, but not being counted as DEGs.


*p_val_adj was calculated using FDR instead of default BH which causes inflated values.

```{r, results='asis', echo=FALSE, message=FALSE, warning=FALSE}
# Print summary
paged_table(deg_summary)
```

# `r colorize('Volcano plot for DEGs for all subclusters', 'blue')` {.tabset}

Not plotting mitochondrial and ribosomal genes. Present in the data but not including them as genes of interest.

```{r,results='asis', echo=FALSE, message=FALSE, warning=FALSE}
library(ggplot2)
library(ggrepel)

# List of CSV files for each subcluster
csv_files <-  list.files("../csvs/tcells/condition/", pattern = "^Tcell_", full.names = TRUE)

# Function to generate volcano plot
volcanoplot <- function(de, subcluster_name) {
  # List of oligo types
  types <- c("HIV")
  
  # Filter out mitochondrial and ribosomal genes
  de <- de %>% filter(!grepl("^MT", gene) & !grepl("^RB[SL]", gene))
  
  de<- subset(de, cluster=="HIV")
  
  de$diffexpressed <- "NO"
  de$diffexpressed[de$avg_log2FC < -0.25 & de$p_val < 0.05] <- "DOWN"
  de$diffexpressed[de$avg_log2FC > 0.25 & de$p_val < 0.05] <- "UP"

  subset_df <- de[complete.cases(de[, "gene"]), ]

  c1 <- subset_df[subset_df[, "avg_log2FC"] > 0 & subset_df[, "p_val"] < 0.05, ]
  c2 <- subset_df[subset_df[, "avg_log2FC"] < 0 & subset_df[, "p_val"] < 0.05, ]
  c1_byavg_log2FC <- c1[order(c1$avg_log2FC, decreasing = TRUE), ]
  c2_byavg_log2FC <- c2[order(c2$avg_log2FC, decreasing = FALSE), ]
  c1_top10 <- head(c1_byavg_log2FC, 10)$gene
  c2_top10 <- head(c2_byavg_log2FC, 10)$gene
  labeled_proteins <- c(c1_top10, c2_top10)
  de$delabel <- ifelse(de$gene %in% labeled_proteins & de$diffexpressed %in% c("UP", "DOWN"), de$X, NA)

  ggplot(data = de, aes(x = avg_log2FC, y = -log10(p_val), col = diffexpressed, label = delabel)) +
    geom_vline(xintercept = c(-0.25, 0.25), col = "gray", linetype = 'dashed') +
    geom_hline(yintercept = -log10(0.05), col = "gray", linetype = 'dashed') + 
    geom_point(data = de %>% filter(diffexpressed == "DOWN"), color = "#00AFBB", size = 0.8, alpha = 0.5) +
    geom_point(data = de %>% filter(diffexpressed == "UP"), color = "#bb0c00", size = 0.8, alpha = 0.5) +
    geom_point(data = de %>% filter(diffexpressed == "NO"), color = "gray", size = 0.8, alpha = 0.5) +
    geom_text_repel(min.segment.length = 0, size = 2, max.overlaps = Inf, show.legend = FALSE) +
    scale_color_manual(values = c("DOWN" = "black", "UP" = "black")) +
    guides(color = guide_legend(title = NULL)) +
    labs(x = expression("Upregulated in CONTROL <- avg log"[2]*"FC -> Upregulated in HIV"), 
         y = expression("-log"[10]*"p-value"),
         title = paste0("DEGs in ", subcluster_name, " HIV vs CONTROL")) +
    theme_bw() +
    theme(plot.title = element_text(hjust = 0.5), axis.text = element_text(size = 16))
}

# Initialize list to store volcano plots
volcano_plots <- list()

# Loop through each CSV file, generate plots, and store them
for (file in csv_files) {
  if (grepl("Tcell_subcluster_\\d+_HIV_vs_Control.csv", file)) {
    subcluster_name <- gsub(".*Tcell_subcluster_(\\d+)_HIV_vs_Control.csv", "Subcluster \\1", file)
  } else if (grepl("Tcell_all_markers_condition.csv", file)) {
    subcluster_name <- "Tcell (Overall)"
  } else {
    subcluster_name <- "Unknown"
  }
  
  de <- read.csv(file)
  
  # Store the plot in the list
  volcano_plots[[subcluster_name]] <- volcanoplot(de, subcluster_name)
}

# Print all volcano plots
for (name in names(volcano_plots)) {
  cat("##", name, "\n")
  print(volcano_plots[[name]])
  cat("\n\n")
}

```

