---
title: "Non-integrated samples"
author: "Saumya"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_depth: 3
    toc_float: true
    theme: cosmo
#    number_sections: true # Adds numbers to headers
    css: styles.css
    code_folding: hide
---


```{r, results = FALSE, message=FALSE, warning=FALSE}
# Load required packages
library(dplyr)
library(Seurat)
library(patchwork)
library(ggplot2)
library(cowplot)
library(rmarkdown)
library(SummarizedExperiment)
library(htmlwidgets)
library(htmltools)
library(shiny)
library(openxlsx)
library(scDblFinder)
library(hdf5r)
library(future)
```

<!-- Information of Used Library -->
## Information of Used R Libraries

The following R libraries were used to process the data and produce this report:

R: `r R.Version()$version.string`,

dplyr: `r packageVersion('dplyr')`, 

Seurat: `r packageVersion('Seurat')`, 

patchwork: `r packageVersion('patchwork')`, 

ggplot2: `r packageVersion('ggplot2')`, 

cowplot: `r packageVersion('cowplot')`,

scDblFinder: `r packageVersion('scDblFinder')`

```{r, results = FALSE, message=FALSE, warning=FALSE}
sampleNames <- c("ABC", "XYZ", "LMN")
data_dir <- ("../../raw_h5")
projectName <- "Seurat"
h5_string <- "_filtered_feature_bc_matrix.h5"
```

```{r,  eval=FALSE, message = FALSE, results= FALSE, warning=FALSE}
# Define filenames and path
filenames <- c(file.path(data_dir, paste0(sampleNames, h5_string)))

# Check if the path for the files is right
filenames

# Create a list to store Seurat objects
seurat_list <- list()
 
# Loop through filenames and create Seurat objects
for (i in 1:length(filenames)) {
  raw_data <- Read10X_h5(filenames[i], use.names = TRUE, unique.features = TRUE)
  seurat_obj <- CreateSeuratObject(counts = raw_data,
                                    project = projectName,
                                   min.cells = 3,
                                   min.features = 200)
# Add sample name to orig.ident
 seurat_obj$orig.ident <- sampleNames[i]

# Store in the list
 seurat_list[[i]] <- seurat_obj
}
 
# Merge all Seurat objects into one
combined_data <- merge(seurat_list[[1]], y = seurat_list[-1], add.cell.ids = sampleNames)
 
combined_data@meta.data$sample_names <- combined_data@meta.data$orig.ident

# Save the combined Seurat object
saveRDS(combined_data, file = "../RDS/data_combined_raw.RDS")
```

```{r, eval= FALSE, results = FALSE, message=FALSE, warning=FALSE}
# Add condition information based on sample names
data <- readRDS("../RDS/data_combined_raw.RDS")

sample_to_group <- c("ABC" = "Control",
                    "XYZ" = "Condition",
                    "LMN" = "Condition")

# Apply the new annotations to your Seurat object
data$condition <- plyr::mapvalues(
  data$sample_names,
   from = names(sample_to_group),
  to = sample_to_group
)

# Check if condition has been mapped appropriately
unique(data$condition)
```


```{r, eval = FALSE, results = FALSE, message=FALSE, warning=FALSE}
saveRDS(data, file = paste("../RDS/data_combined_raw.RDS"))
```


```{r, eval = FALSE, results = FALSE, message=FALSE, warning=FALSE}
# Standard Seurat workflow
data <- readRDS("../RDS/data_combined_raw.RDS")

# Standard normalisation workflow
data <- NormalizeData(data, normalization.method = "LogNormalize", scale.factor = 10000)
data <- FindVariableFeatures(data, selection.method = "vst", nfeatures = 2000)
data <- ScaleData(data, verbose = TRUE)

#dimension reduction
data <- RunPCA(data, npcs = 50, verbose = FALSE)

# Check how many PCs to use
ElbowPlot(data, ndims = 50)

# Use appropriate number of PCs
data <- RunUMAP(data, reduction = "pca", dims = 1:20)
```

```{r, eval = FALSE, results = FALSE, message=FALSE, warning=FALSE}
saveRDS(data, file = paste("../RDS/data_combined_raw_umap.RDS"))
```

# Non-integrated Samples before QC {.tabset}

```{r, results = 'asis', message=FALSE, warning=FALSE}
data <- readRDS("../RDS/data_combined_raw_umap.RDS")

cat("##", "Samples", " \n")
DimPlot(data, reduction = "umap", group.by = "sample_names", raster = FALSE) + ggtitle ("Samples before integration and QC") + theme(legend.position = "none")
cat(" \n \n")

cat("##", "Condition", " \n")
DimPlot(data, reduction = "umap", group.by = "condition", raster = FALSE) + ggtitle ("Condition before integration and QC")
cat(" \n \n")
```


# QC: Checking mitochondrial and ribosomal gene percentages {.tabset}

Cells with > 10% mitochondrial percentage and nFeature_RNA <200 OR >2500 are generally filtered. 

Note: nFeature_RNA is the total number of genes detected in each cell. It counts how many genes show a non-zero expression level for a given cell, and <200 is a low quality cell, and >2500 is assumed to be a doublet due to higher than normal expression levels, but data varies. Make sure to check your data and determine appropriate cutoffs to not lose real biological information.

```{r, results = FALSE, message=FALSE, warning=FALSE}
# Check mitochondrial gene names
gene_names <- rownames(data)
mt_genes <- grep("^MT-", gene_names, value = TRUE)
print(mt_genes)

# Calculate mitochondrial gene percentage
data[["percent.mt"]] <- PercentageFeatureSet(data, pattern = "^MT-")
 
# Add new column with mitochondrial percentage information
data$mt_label <- factor(ifelse(data$percent.mt > 20, "Above 20%", "Below 20%"))
 

# Check ribosomal gene names
gene_names <- rownames(data)
rp_genes <- grep("^RP[SL]", gene_names, value = TRUE)
print(rp_genes)

# Calculate ribosomal gene percentage
data[["percent.rp"]] <- PercentageFeatureSet(data, pattern = "^RP[SL]")

# Add new column with ribosomal percentage information
data$rp_label <- factor(ifelse(data$percent.rp > 10, "Above 10%", "Below 10%"))
```

```{r, eval = FALSE, results = FALSE, message=FALSE, warning=FALSE}
# Save RDS with mitochondrial and ribosomal percentages
saveRDS(data, file = paste("../RDS/data_combined_raw_umap.RDS"))
```

```{r, results = 'asis', message=FALSE, warning=FALSE}
data <- readRDS("../RDS/data_combined_raw_umap.RDS")

cat("##", "Mitochondrial %", " \n")
DimPlot(data, reduction = "umap", group.by = "mt_label", raster = FALSE) + ggtitle ("Percentage of Mitochondrial Genes")
VlnPlot(data, features = c("percent.mt"), pt.size = 0.0, group.by = "sample_names") + ggtitle ("Percentage of Mitochondrial Genes") + theme(legend.position = "none") + coord_cartesian(ylim = c(0, 25))
cat(" \n \n")

cat("##", "Ribosomal %", " \n")
DimPlot(data, reduction = "umap", group.by = "rp_label", raster = FALSE) + ggtitle ("Percentage of Ribosomal Genes")
VlnPlot(data, features = c("percent.rp"), pt.size = 0.0, group.by = "sample_names") + ggtitle ("Percentage of Ribosomal Genes") + theme(legend.position = "none") + coord_cartesian(ylim = c(0, 25))
cat(" \n \n")
```



```{r, results = FALSE, eval = FALSE, message=FALSE, warning=FALSE
# Plot and save integrated umap
pdf(paste("../pdfs/non_integrated_mitochondrial_pct.pdf"), width=16, height=10, useDingbats = FALSE)
DimPlot(data, reduction = "umap", group.by = "mt_label", raster = FALSE) + ggtitle ("Percentage of Mitochondrial Genes")
VlnPlot(data, features = c("percent.mt"), pt.size = 0.0, group.by = "sample") + ggtitle ("Percentage of Mitochondrial Genes") + theme(legend.position = "none") + coord_cartesian(ylim = c(0, 25))
dev.off()
```

```{r, results = FALSE, message=FALSE, warning=FALSE}
# Filter out cells with >20% mitochondrial gene exp. Change according to data
data_filtered <- subset(data, subset = nFeature_RNA > 200 & nFeature_RNA < 2500 & percent.mt < 20)
```

```{r, eval = FALSE, results = FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# Save RDS post filtering
saveRDS(data_filtered, file = paste("../RDS/data_filtered_umap.RDS"))
```

```{r, results=FALSE, message=FALSE, warning=FALSE}
# Read pre and post filtering RDS files to compute statistics on how many cells are being filtered

data_filtered <- readRDS("../RDS/data_filtered_umap.RDS")
data <- readRDS("../RDS/data_combined_raw_umap.RDS")

# Get total and filtered cell counts 
total_cells <- length(Cells(data))
retained_cells <- length(Cells(data_filtered))
filtered_cells <- total_cells - retained_cells  

# Calculate percentages
retained_percentage <- (retained_cells / total_cells) * 100
filtered_percentage <- (filtered_cells / total_cells) * 100

# Count cells with NA values in filtering columns
na_filtered <- sum(is.na(data$nFeature_RNA) | is.na(data$percent.mt))

# Count cells removed due to BOTH filters (overlap)
overlap_filtered <- sum((data$nFeature_RNA < 200 | data$nFeature_RNA > 2500) & (data$percent.mt > 20), na.rm = TRUE)

# Count cells removed for nFeature_RNA and percent.mt, adjusting for overlap
nfeature_count <- sum(data$nFeature_RNA < 200 | data$nFeature_RNA > 2500, na.rm = TRUE) 
mitochondrial_cells <- sum(data$percent.mt > 20, na.rm = TRUE)

# Correct the total filtered count by adding NAs
filtered_cells_corrected <- nfeature_count + mitochondrial_cells - overlap_filtered + na_filtered

# Summary table
summary_table <- data.frame(
  "Category" = c("Total Cells", "Retained Cells", "Total Filtered Cells", "nFeature_RNA Cells", "Mitochondrial Cells"),
  "Raw Count" = c(total_cells, retained_cells, filtered_cells, nfeature_count, mitochondrial_cells),
  "Percentage" = c(
    100, 
    retained_percentage, 
    (filtered_cells / total_cells) * 100, 
    (nfeature_count / total_cells) * 100, 
    (mitochondrial_cells / total_cells) * 100
  )
)

# Format to remove scientific notation
summary_table$Percentage <- format(summary_table$Percentage, scientific = FALSE)
```

# Cell Count Summary

```{r, results='asis', message=FALSE, warning=FALSE}
# Display summary table
paged_table(summary_table)
```


# Non-integrated Samples after mitochondrial and nFeature RNA Filtering {.tabset}

```{r, results = 'asis', echo=FALSE, message=FALSE, warning=FALSE}

data_filtered <- readRDS("../RDS/data_filtered_umap.RDS")

cat("##", "Samples", " \n")
DimPlot(data, reduction = "umap", group.by = "sample_names", raster = FALSE) + ggtitle ("UMAP post Mitochondrial Filtering") + theme(legend.position = "none")

cat(" \n \n")
cat("##", "Condition", " \n")
DimPlot(data, reduction = "umap", group.by = "condition", raster = FALSE) + ggtitle ("UMAP post Mitochondrial Filtering")
cat(" \n \n")
```


# Doublet Detection

```{r, eval=FALSE, results = FALSE, message=FALSE, warning=FALSE}
# Doublet detection using scDblfinder)
data_filtered <- readRDS("../RDS/data_filtered_umap.RDS")
DefaultAssay(data_filtered) <- "RNA"
set.seed(100)

sce <- data_filtered
sce <- as.SingleCellExperiment(sce)
sce <- scDblFinder(sce)
table(sce$scDblFinder.class)
sce@colData@listData %>% as.data.frame() %>% head()
```

```{r, eval= FALSE, results = FALSE, message=FALSE, warning=FALSE}
meta_scdblfinder <- sce@colData@listData %>% as.data.frame() %>%
  dplyr::select(starts_with('scDblFinder')) # 'scDblFinder.class')
head(meta_scdblfinder)

# Explore results and add to seurat object
meta_scdblfinder <- sce@colData@listData %>% as.data.frame() %>%
  dplyr::select(starts_with('scDblFinder')) # 'scDblFinder.class')
head(meta_scdblfinder)

rownames(meta_scdblfinder) <- sce@colData@rownames
head(meta_scdblfinder)
data_filtered<- AddMetaData(object = data_filtered, metadata = meta_scdblfinder %>% dplyr::select('scDblFinder.class'))
```

```{r, eval = FALSE, results = FALSE, message=FALSE, warning=FALSE}
# Save RDS with doublet information
saveRDS(data_filtered, file = paste("../RDS/data_filtered_umap.RDS"))
```

```{r, results = FALSE,  message=FALSE, warning=FALSE}
# Read both RDS files to calculate metrics on how many cells would be retained after doublet filtering
data_filtered <- readRDS("../RDS/data_filtered_umap.RDS")
data <- readRDS("../RDS/data_combined_raw_umap.RDS)

# Get the total number of cells in the dataset
total_cells <- length(Cells(data_filtered))

# Get the count of cells retained after filtering
singlet_count <- sum(data_filtered$scDblFinder.class == "singlet")

# Calculate the number of mitochondrial cells 
doublet_count <- sum(data_filtered$scDblFinder.class == "doublet")

# Calculate percentages
singlet_percentage <- (singlet_count / total_cells) * 100
doublet_percentage <- (doublet_count / total_cells) * 100

# Create a summary table
doublet_summary <- data.frame(
  "Category" = c("Total Cells", "Singlet Cells", "Doublet Cells"),
  "Raw Count" = c(total_cells, singlet_count, doublet_count),
  "Percentage" = c(100, singlet_percentage, doublet_percentage)
)
```

## Doublet Count Summary {.tabset}

After mitchondrial >20%  and nFeature_RNA (<200 OR >2500) filtering.

```{r, results='asis', message=FALSE, warning=FALSE}
# Display the summary table
cat("###", "Doublet Summary", " \n")
paged_table(doublet_summary)
cat(" \n \n")

# Visualize with the annotated cell types
cat("###", "Doublet Dimplot", " \n")
DimPlot(data_filtered, reduction = "umap", group.by="scDblFinder.class") + ggtitle("Doublet Distribution")
cat(" \n \n")
```

```{r, results = FALSE, message=FALSE, warning=FALSE}
# Filter out cells that are doublets according to scDblFinder and save only singlets
seu_dblt <- subset(data.combined, scDblFinder.class == 'singlet')

saveRDS(seu_dblt, "../RDS/data_dblt_filtered_umap.RDS")
```


# Cell Count Distribution {.tabset}

```{r, results=FALSE, message=FALSE, warning=FALSE}
# Calculate cell distribution sample wise, raw counts and percentages

# Load filtered dataset
data_filtered <- readRDS("../RDS/data_dblt_filtered_umap.RDS")

# Count number of cells per sample
cell_counts <- table(data_filtered$sample_names)

# Convert to data frame
df_counts <- as.data.frame(cell_counts)
colnames(df_counts) <- c("Sample", "Cell_Count")

# Calculate percentage distribution (ensuring it sums to 100)
df_counts$Percentage <- (df_counts$Cell_Count / sum(df_counts$Cell_Count)) * 100

# Bar plot for percentage distribution with y-axis fixed from 0 to 100
p1 <- ggplot(df_counts, aes(x = Sample, y = Percentage, fill = Sample)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(
    title = "Cell Percentage Distribution Across Samples",
    x = "Sample",
    y = "Percentage of Total Cells",
    fill = "Sample"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none")

# Bar plot for raw counts
p2 <- ggplot(df_counts, aes(x = Sample, y = Cell_Count, fill = Sample)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(
    title = "Raw Cell Count Distribution Across Samples",
    x = "Sample",
    y = "Number of Cells",
    fill = "Sample"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none")
```


```{r, results='asis', message=FALSE, warning=FALSE}
cat("##", "Raw Count Distribution", " \n")
print(p2)
cat ('\n \n')

cat("##", "Percentage Distribution", " \n")
print(p1)
cat ('\n \n')
```
