---
title: "Samples Clustering"
author: "Saumya"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_depth: 3
    toc_float: true
    theme: cosmo
#    number_sections: true # Adds numbers to headers
    css: styles.css
    code_folding: hide
---

```{r, results = FALSE, message=FALSE, warning=FALSE}
# Load packages
library(dplyr)
library(Seurat)
library(patchwork)
library(ggplot2)
library(cowplot)
library(rmarkdown)
library(SummarizedExperiment)
library(htmlwidgets)
library(htmltools)
library(shiny)
library(openxlsx)
library(scDblFinder)
library(hdf5r)
library(harmony)
library(DT)
library(SeuratWrappers)
```

<!-- Information of Used Library -->
## Information of Used R Libraries 

The following R libraries were used to process the data and produce this report.

R: `r R.Version()$version.string`,

dplyr: `r packageVersion('dplyr')`, 

Seurat: `r packageVersion('Seurat')`, 

patchwork: `r packageVersion('patchwork')`, 

ggplot2: `r packageVersion('ggplot2')`, 

cowplot: `r packageVersion('cowplot')`,

scDblFinder: `r packageVersion('scDblFinder')`

```{r, results = FALSE, message=FALSE, warning=FALSE}
sampleNames <- c("ABC", "XYZ", "LMN")
data_dir <- ("../../raw_h5")
projectName <- "Seurat"
h5_string <- "_filtered_feature_bc_matrix.h5"
```

# Clustering for Samples

```{r,results='asis', message=FALSE, warning=FALSE}
# UMAP before clustering

data.combined <- readRDS("../RDS/data_Integrated_umap.RDS")
Idents(data.combined) <- "sample_names"
DimPlot(data.combined, reduction = "umap") + ggtitle ("Before Clustering") + theme(legend.position = "none")
```

```{r, eval = FALSE, message = FALSE, results= FALSE, warning=FALSE}
# Do clustering

data.combined <- FindNeighbors(data.combined, reduction = "harmony", dims = 1:20)
data.combined <- FindClusters(data.combined)
```

```{r, eval = FALSE, message = FALSE, results= FALSE, warning=FALSE}
# Save clustering data
saveRDS(data.combined, file = paste("../RDS/data_umap_33_clusters.RDS") )
```

```{r,results='asis', echo=FALSE, message=FALSE, warning=FALSE}
# UMAP after clustering
data.combined <- readRDS("../RDS/data_umap_33_clusters.RDS")
Idents(data.combined) <- "seurat_clusters"

# Clustering UMAP
DimPlot(data.combined, reduction = "umap", label = TRUE, repel = TRUE, raster=FALSE) + ggtitle("Seurat Clusters")
```

# Clustering Step 2: Find Markers

Find markers for every cluster compared to all remaining cells, report both positive and negative ones. Showing the top 20 markers (selected based on the largest values of avg_log2FC or pct.ratio) and also all markers for each cluster on the following data table. The results data frame has the following columns:

* cluster: cluster label.
* gene: gene name
* p_val: p value (unadjusted) can take any value between 0 and 1. Values close to 0 indicate that the observed gene expression difference between the cluster and the rest clusters is unlikely to be due to chance, whereas a p value close to 1 suggests no difference between the cluster and the rest clusters other than due to chance.
* avg_log2FC: log fold-change of the average expression between the cluster and the rest clusters. Positive values indicate that the gene is more highly expressed in the cluster.
* pct.1: The percentage of cells where the gene is detected in the cluster.
* pct.2: The percentage of cells where the gene is detected in the rest clusters.
* p_val_adj: Adjusted p value, based on Bonferroni correction using all genes in the dataset.
* pct.ratio: The ratio between pct.1 and pct.2.


```{r, eval = FALSE, message=FALSE, warning=FALSE}
# Find All Markers for Seurat clusters. Idents can be set to condition/site or any other column as required

data.combined <- readRDS("../RDS/data_umap_33_clusters.RDS")
DefaultAssay(data.combined) <- "RNA"
Idents(data.combined) <- "seurat_clusters"

data.combined <- JoinLayers(data.combined)

data.markers <- SeuratWrappers::RunPrestoAll(data.combined, only.pos = FALSE, min.pct = 0.25, logfc.threshold = 0.25)

# Save markers
write.csv(data.markers, file='../csvs/data_all_markers_33_clusters.csv')
```


```{r, results = FALSE, message=FALSE, warning=FALSE}
# Read Markers .csv
data.markers = read.csv('../csvs/data_all_markers_33_clusters.csv')
data.markers$pct.diff = data.markers$pct.1 - data.markers$pct.2
data.markers$pct.ratio = data.markers$pct.1 / data.markers$pct.2
avglog_p = data.markers %>% group_by(cluster) %>% slice_max(n = 20, order_by = avg_log2FC)
avglog_n = data.markers %>% group_by(cluster) %>% slice_min(n = 20, order_by = avg_log2FC)
pctdiff_p = data.markers %>% group_by(cluster) %>% slice_max(n = 20, order_by = pct.diff)
pctdiff_n = data.markers %>% group_by(cluster) %>% slice_min(n = 20, order_by = pct.diff)
pctratio_p = data.markers %>% group_by(cluster) %>% slice_max(n = 20, order_by = pct.ratio)
pctratio_n = data.markers %>% group_by(cluster) %>% slice_min(n = 20, order_by = pct.ratio)
```


```{r, eval = FALSE, results = FALSE, message=FALSE, warning=FALSE}
# Store only positive markers in a separate csv file
positive.markers = read.csv('../csvs/data_all_markers_33_clusters.csv')
positive.markers$pct.diff = positive.markers$pct.1 - positive.markers$pct.2
positive.markers$pct.ratio = positive.markers$pct.1 / positive.markers$pct.2

# Keep positive markers only
positive.markers <- positive.markers[positive.markers$pct.diff > 0.0,]
positive.markers = positive.markers %>% group_by(cluster) %>% slice_max(n = 200, order_by = pct.diff)

write.csv(positive.markers, file='../csvs/data_pos_markers_33_clusters.csv')
```

## Positive markers {.tabset}

While downloading tables please make sure to change filename, otherwise it might overwrite if downloading multiple tables. Make sure to include ".csv" after changing filename to retain format.

### Top 20 Markers (avg_log2FC)

```{r,results='asis', message=FALSE, warning=FALSE}
# Create a table with a CSV download option
datatable(
  avglog_p, 
  extensions = 'Buttons', 
  options = list(
    dom = 'Bfrtip',
    buttons = list('csv'),
    filter = "top"  
  )
)
```

### Top 20 Markers (pctdiff)

```{r,results='asis', message=FALSE, warning=FALSE}
datatable(
  pctdiff_p, 
  extensions = 'Buttons', 
  options = list(
    dom = 'Bfrtip',
    buttons = list('csv'),
    filter = "top"
  )
)
```

### Top 20 Markers (pct.ratio)

```{r,results='asis', message=FALSE, warning=FALSE}
datatable(
  pctratio_p, 
  extensions = 'Buttons', 
  options = list(
    dom = 'Bfrtip',
    buttons = list('csv', 'search')
  )
)
```


## Negative markers {.tabset}

While downloading tables please make sure to change filename, otherwise it might overwrite if downloading multiple tables. Make sure to include ".csv" after changing filename to retain format.

### Top 20 Markers (avg_log2FC)

```{r,results='asis', message=FALSE, warning=FALSE}
# Create a table with a CSV download option
datatable(
  avglog_n, 
  extensions = 'Buttons', 
  options = list(
    dom = 'Bfrtip',
    buttons = list('csv'),
    filter = "top"  
  )
)
```

### Top 20 Markers (pctdiff)

```{r,results='asis', message=FALSE, warning=FALSE}
datatable(
  pctdiff_n, 
  extensions = 'Buttons', 
  options = list(
    dom = 'Bfrtip',
    buttons = list('csv'),
    filter = "top"
  )
)
```

### Top 20 Markers (pct.ratio)

```{r,results='asis', message=FALSE, warning=FALSE}
datatable(
  pctratio_n, 
  extensions = 'Buttons', 
  options = list(
    dom = 'Bfrtip',
    buttons = list('csv', 'search')
  )
)
```

## All markers

While downloading tables please make sure to change filename, otherwise it might overwrite if downloading multiple tables. Make sure to include ".csv" after changing filename to retain format.

```{r,results='asis', message=FALSE, warning=FALSE}
# Create a table with a CSV download option
datatable(
  data.markers, 
  extensions = 'Buttons', 
  options = list(
    dom = 'Bfrtip',
    buttons = list('csv')
  )
)
```


```{r, results = FALSE, message=FALSE, warning=FALSE}
data.combined <- readRDS("../RDS/data_umap_33_clusters.RDS")

# Annotate all clusters appropriately
cluster_annotations <- c("0" = "Plasma cells",
                         "1" = "Plasma cells")

# Apply the new annotations to your Seurat object
data.combined$cluster_label <- plyr::mapvalues(
 Idents(data.combined),
    from = names(cluster_annotations),
  to = cluster_annotations
)

# Update the identities in the Seurat object
Idents(data.combined) <- data.combined$cluster_label
 
saveRDS(data.combined, "../RDS/data_umap_33_clusters.RDS")
```


# Cell Count Distribution in Clusters {.tabset}

```{r, results = FALSE, message=FALSE, warning=FALSE}
# Calculate cell count distribution cluster wise
data.combined <- readRDS("../RDS/data_umap_33_clusters.RDS")

# Count number of cells per sample
cell_counts <- table(data.combined$seurat_clusters)

# Convert to data frame
df_counts <- as.data.frame(cell_counts)
colnames(df_counts) <- c("Cluster", "Cell_Count")

# Calculate percentage distribution
df_counts$Percentage <- (df_counts$Cell_Count / sum(df_counts$Cell_Count)) * 100

# Bar plot for percentage distribution with y-axis
p1 <- ggplot(df_counts, aes(x = Cluster, y = Percentage, fill = Cluster)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(
    title = "Cell Percentage Distribution Across Clusters",
    x = "Cluster",
    y = "Percentage of Total Cells",
    fill = "Cluster"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none")

# Bar plot for raw counts
p2 <- ggplot(df_counts, aes(x = Cluster, y = Cell_Count, fill = Cluster)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(
    title = "Raw Cell Count Distribution Across Clusters",
    x = "Cluster",
    y = "Number of Cells",
    fill = "Cluster"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none")
```


```{r, results='asis', message=FALSE, warning=FALSE}
cat("##", "Raw Count Distribution", " \n")
print(p2)
cat ('\n \n')

cat("##", "Percentage Distribution", " \n")
print(p1)
cat ('\n \n')
```


```{r, results = FALSE, message=FALSE, warning=FALSE}
# Extract raw counts for the one cluster of interest if need be (e.g, cluster 24)
24_cells <- WhichCells(data.combined, ident = "24")
 
# Get the raw count of neutrophil cells
24_count <- length(24_cells)
 
# Get the total number of cells in the dataset
total_cells <- length(Cells(data.combined))
 
# Calculate the percentage of neutrophils in the dataset
24_percentage <- (24_count / total_cells) * 100
 
# Create a summary table
summary_table <- data.frame(
   "Cell Type" = c("24", "Total Cells"),
  "Raw Count" = c(24_count, total_cells),
   "Percentage" = c(24_percentage, 100)
)
```

# Cell Count for Cluster 24 {.tabset}

```{r, results='asis', message=FALSE, warning=FALSE}
# Display the summary table
paged_table(summary_table)
```

