---
title: "Subset Clustering"
author: "Saumya"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_depth: 3
    toc_float: true
    theme: cosmo
#    number_sections: true # Adds numbers to headers
    css: styles.css
    code_folding: hide
---

```{r, results = FALSE, message=FALSE, warning=FALSE}
# Load packages
library(dplyr)
library(Seurat)
library(patchwork)
library(ggplot2)
library(cowplot)
library(rmarkdown)
library(SummarizedExperiment)
library(htmlwidgets)
library(htmltools)
library(shiny)
library(openxlsx)
library(scDblFinder)
library(hdf5r)
library(harmony))
library(DT)
```

<!-- Information of Used Library -->
## Information of Used R Libraries

The following R libraries were used to process the data and produce this report.

R: `r R.Version()$version.string`,

dplyr: `r packageVersion('dplyr')`, 

Seurat: `r packageVersion('Seurat')`, 

patchwork: `r packageVersion('patchwork')`, 

ggplot2: `r packageVersion('ggplot2')`, 

cowplot: `r packageVersion('cowplot')`,

scDblFinder: `r packageVersion('scDblFinder')`

```{r, results = FALSE, message=FALSE, warning=FALSE}
sampleNames <- c("ABC", "XYZ", "LMN")
data_dir <- ("../../raw_h5")
projectName <- "Seurat"
h5_string <- "_filtered_feature_bc_matrix.h5"
```

```{r, results = FALSE, eval= FALSE, message=FALSE, warning=FALSE}
data.combined <- readRDS("../RDS/data_umap_29_clusters.RDS")
Idents(data.combined) <- "seurat_clusters"
subset_data <- subset(data.combined, idents = c("14","15"))
```

```{r, eval=FALSE, results = FALSE, message=FALSE, warning=FALSE}
subset_data <- FindVariableFeatures(subset_data, selection.method = "vst", nfeatures = 2000)
subset_data <- ScaleData(subset_data, verbose = FALSE)
 
#dimension reduction
subset_data <- RunPCA(subset_data, npcs = 50, verbose = FALSE)

# Check how many PCs to use
ElbowPlot(subset_data, ndims = 50)

subset_data <- RunUMAP(subset_data, reduction = "harmony", dims = 1:50)

# Update the identities in the Seurat object
Idents(subset_data) <- subset_data$cluster_label
```

```{r, eval = FALSE, results = FALSE, message=FALSE, warning=FALSE}
# Save subsetted data
saveRDS(subset_data, file = paste("../RDS/subset_umap.RDS"))
```

# Subset before clustering

```{r,results='asis', message=FALSE, warning=FALSE}
# Subset UMAP before clustering. Sometimes subsets need to be reintegrated. Check if this is necessary by using DimPlot with samples as well as condition
subset_data <- readRDS("../RDS/subset_umap.RDS")
Idents(subset_data) <- "cluster_label"
DimPlot(subset_data, reduction = "umap") + ggtitle ("Subset before clustering") 
```

```{r, results=FALSE, message=FALSE, warning=FALSE}
# Extract counts for subset data

data.combined <- readRDS("../RDS/data_umap_29_clusters.RDS")
subset_data <- readRDS("../RDS/subset_umap.RDS")

# Get the raw count of neutrophil cells
subset_count <- length(Cells(subset_data))

# Get the total number of cells in the dataset
total_cells <- length(Cells(data.combined))

# Calculate the percentage of neutrophils in the dataset
subset_percentage <- (subset_count / total_cells) * 100

# Create a summary table
summary_table <- data.frame(
  "Cell Type" = c("Subset Cells", "Total Cells"),
  "Raw Count" = c(subset_count, total_cells),
  "Percentage" = c(subset_percentage, 100)
)
```

# Cell Count for Subset Cells {.tabset}
```{r, results='asis', message=FALSE, warning=FALSE}
# Display the summary table
paged_table(summary_table)
```


# Sub-clustering for Subset cells

```{r, eval = FALSE, message = FALSE, results= FALSE, warning=FALSE}
# Do clustering
subset_data <- readRDS("../RDS/subset_umap.RDS")
subset_data <- FindNeighbors(subset_data, reduction = "harmony", dims = 1:50)
subset_data <- FindClusters(subset_data, resolution = 0.5)
```

```{r, eval = FALSE, message = FALSE, results= FALSE, warning=FALSE}
# Save clustering data
saveRDS(subset_data, file = paste("../RDS/subset_umap_12_clusters.RDS") )
```

```{r,results='asis', message=FALSE, warning=FALSE}
subset_data <- readRDS("../RDS/subset_umap_12_clusters.RDS")
Idents(subset_data) <- "seurat_clusters"

# Check clustering UMAP
DimPlot(subset_data, reduction = "umap", label = TRUE, repel = TRUE, raster=FALSE) + ggtitle("Seurat Clusters")
```

# DEGS for 12 Subset sub-clusters

```{r, eval = FALSE, message=FALSE, warning=FALSE}
# Find All Markers
subset_data <- readRDS("../RDS/subset_umap_12_clusters.RDS")
DefaultAssay(subset_data) <- "RNA"
Idents(subset_data) <- "seurat_clusters"

subset_data <- JoinLayers(subset_data)

data.markers <- FindAllMarkers(subset_data, only.pos = FALSE, min.pct = 0.25, logfc.threshold = 0.25)
data.markers$p_val_adj <- p.adjust(data.markers$p_val, method = "fdr")

# Save markers
write.csv(data.markers, file='../csvs/subset_all_markers_12_clusters.csv')
```

```{r, results = FALSE, message=FALSE, warning=FALSE}
# Read Markers .csv
data.markers = read.csv('../csvs/subset_all_markers_12_clusters.csv')
data.markers$pct.diff = data.markers$pct.1 - data.markers$pct.2
data.markers$pct.ratio = data.markers$pct.1 / data.markers$pct.2
avglog_p = data.markers %>% group_by(cluster) %>% slice_max(n = 20, order_by = avg_log2FC)
avglog_n = data.markers %>% group_by(cluster) %>% slice_min(n = 20, order_by = avg_log2FC)
pctdiff_p = data.markers %>% group_by(cluster) %>% slice_max(n = 20, order_by = pct.diff)
pctdiff_n = data.markers %>% group_by(cluster) %>% slice_min(n = 20, order_by = pct.diff)
pctratio_p = data.markers %>% group_by(cluster) %>% slice_max(n = 20, order_by = pct.ratio)
pctratio_n = data.markers %>% group_by(cluster) %>% slice_min(n = 20, order_by = pct.ratio)
```

```{r, eval = FALSE, results = FALSE, message=FALSE, warning=FALSE}
# Generate positive marker DEG list
positive.markers = read.csv('../csvs/subset_all_markers_12_clusters.csv')

positive.markers$pct.diff = positive.markers$pct.1 - positive.markers$pct.2
positive.markers$pct.ratio = positive.markers$pct.1 / positive.markers$pct.2

# Keep positive markers only
positive.markers <- positive.markers[positive.markers$pct.diff > 0.0,]
positive.markers = positive.markers %>% group_by(cluster) %>% slice_max(n = 200, order_by = pct.diff)

write.csv(positive.markers, file='../csvs/subset_pos_markers_12_clusters.csv')
```

## Positive markers {.tabset}

While downloading tables please make sure to change filename, otherwise it might overwrite if downloading multiple tables. Make sure to include ".csv" after changing filename to retain format.

### Top 20 Markers (avg_log2FC)

```{r,results='asis', message=FALSE, warning=FALSE}
# Create a table with a CSV download option
datatable(
  avglog_p, 
  extensions = 'Buttons', 
  options = list(
    dom = 'Bfrtip',
    buttons = list('csv'),
    filter = "top"  # Ensures a search box appears
  )
)
```

### Top 20 Markers (pctdiff)

```{r,results='asis',message=FALSE, warning=FALSE}
datatable(
  pctdiff_p, 
  extensions = 'Buttons', 
  options = list(
    dom = 'Bfrtip',
    buttons = list('csv'),
    filter = "top"
  )
)
```

### Top 20 Markers (pct.ratio)

```{r,results='asis', message=FALSE, warning=FALSE}
datatable(
  pctratio_p, 
  extensions = 'Buttons', 
  options = list(
    dom = 'Bfrtip',
    buttons = list('csv', 'search')
  )
)
```

## Negative markers {.tabset}

While downloading tables please make sure to change filename, otherwise it might overwrite if downloading multiple tables. Make sure to include ".csv" after changing filename to retain format.

### Top 20 Markers (avg_log2FC)

```{r,results='asis', message=FALSE, warning=FALSE}
# Create a table with a CSV download option
datatable(
  avglog_n, 
  extensions = 'Buttons', 
  options = list(
    dom = 'Bfrtip',
    buttons = list('csv'),
    filter = "top"  # Ensures a search box appears
  )
)
```

### Top 20 Markers (pctdiff)

```{r,results='asis', message=FALSE, warning=FALSE}
datatable(
  pctdiff_n, 
  extensions = 'Buttons', 
  options = list(
    dom = 'Bfrtip',
    buttons = list('csv'),
    filter = "top"
  )
)
```

### Top 20 Markers (pct.ratio)

```{r,results='asis', message=FALSE, warning=FALSE}
datatable(
  pctratio_n, 
  extensions = 'Buttons', 
  options = list(
    dom = 'Bfrtip',
    buttons = list('csv', 'search')
  )
)
```


## All markers

While downloading tables please make sure to change filename, otherwise it might overwrite if downloading multiple tables. Make sure to include ".csv" after changing filename to retain format.

```{r,results='asis', message=FALSE, warning=FALSE}
# Create a table with a CSV download option
datatable(
  data.markers, 
  extensions = 'Buttons', 
  options = list(
    dom = 'Bfrtip',
    buttons = list('csv')
  )
)
```

```{r, results = FALSE, eval = FALSE, message=FALSE, warning=FALSE}
subset_data <- readRDS("../RDS/subset_umap_12_clusters.RDS")

cluster_annotations <- c("0" = "Plasma",
                         "1" = "Plasma")

# Apply the new annotations to your Seurat object
subset_data$cluster_label <- plyr::mapvalues(
  Idents(subset_data),
  from = names(cluster_annotations),
   to = cluster_annotations
)

# Update the identities in the Seurat object
Idents(subset_data) <- subset_data$cluster_label
 
saveRDS(subset_data, "../RDS/subset_umap_12_clusters.RDS")
```

# Annotated subset clusters

```{r,results='asis', message=FALSE, warning=FALSE, fig.width=12, fig.height=8}
subset_data <- readRDS("../RDS/subset_umap_12_clusters.RDS")
Idents(subset_data) <- "cluster_label"
# Check clustering UMAP
DimPlot(subset_data, reduction = "umap", label = TRUE, repel = TRUE, raster=FALSE) + ggtitle("Seurat Clusters")
```

# Condition Wise DEGs

```{r, eval = FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# Find All Markers
subset_data <- readRDS("../RDS/subset_umap_12_clusters.RDS")
DefaultAssay(subset_data) <- "RNA"
Idents(subset_data) <- "condition"

subset_data <- JoinLayers(subset_data)

data.markers <- FindAllMarkers(subset_data, only.pos = FALSE, min.pct = 0.25, logfc.threshold = 0.25)
data.markers$p_val_adj <- p.adjust(data.markers$p_val, method = "fdr")

# Save markers
write.csv(data.markers, file='../csvs/subset_all_markers_condition.csv')
```


```{r, results = FALSE, message=FALSE, warning=FALSE}
# Read Markers .csv
data.markers = read.csv('../csvs/subset_all_markers_condition.csv')
data.markers$pct.diff = data.markers$pct.1 - data.markers$pct.2
data.markers$pct.ratio = data.markers$pct.1 / data.markers$pct.2
avglog_p = data.markers %>% group_by(cluster) %>% slice_max(n = 20, order_by = avg_log2FC)
avglog_n = data.markers %>% group_by(cluster) %>% slice_min(n = 20, order_by = avg_log2FC)
pctdiff_p = data.markers %>% group_by(cluster) %>% slice_max(n = 20, order_by = pct.diff)
pctdiff_n = data.markers %>% group_by(cluster) %>% slice_min(n = 20, order_by = pct.diff)
pctratio_p = data.markers %>% group_by(cluster) %>% slice_max(n = 20, order_by = pct.ratio)
pctratio_n = data.markers %>% group_by(cluster) %>% slice_min(n = 20, order_by = pct.ratio)
```


```{r, eval = FALSE, results = FALSE, message=FALSE, warning=FALSE}

positive.markers = read.csv('../csvs/subset_pos_markers_condition.csv')

positive.markers$pct.diff = positive.markers$pct.1 - positive.markers$pct.2
positive.markers$pct.ratio = positive.markers$pct.1 / positive.markers$pct.2

# Keep positive markers only
positive.markers <- positive.markers[positive.markers$pct.diff > 0.0,]
positive.markers = positive.markers %>% group_by(cluster) %>% slice_max(n = 200, order_by = pct.diff)

write.csv(positive.markers, file='../csvs/subset_pos_markers_condition.csv')
```

## Positive markers {.tabset}

While downloading tables please make sure to change filename, otherwise it might overwrite if downloading multiple tables. Make sure to include ".csv" after changing filename to retain format.

### Top 20 Markers (avg_log2FC)

```{r,results='asis', message=FALSE, warning=FALSE}
# Create a table with a CSV download option
datatable(
  avglog_p, 
  extensions = 'Buttons', 
  options = list(
    dom = 'Bfrtip',
    buttons = list('csv'),
    filter = "top"  # Ensures a search box appears
  )
)
```

### Top 20 Markers (pctdiff)

```{r,results='asis', message=FALSE, warning=FALSE}
datatable(
  pctdiff_p, 
  extensions = 'Buttons', 
  options = list(
    dom = 'Bfrtip',
    buttons = list('csv'),
    filter = "top"
  )
)
```

### Top 20 Markers (pct.ratio)

```{r,results='asis', message=FALSE, warning=FALSE}
datatable(
  pctratio_p, 
  extensions = 'Buttons', 
  options = list(
    dom = 'Bfrtip',
    buttons = list('csv', 'search')
  )
)
```

## Negative markers {.tabset}
While downloading tables please make sure to change filename, otherwise it might overwrite if downloading multiple tables. Make sure to include ".csv" after changing filename to retain format.

### Top 20 Markers (avg_log2FC)

```{r,results='asis', message=FALSE, warning=FALSE}
# Create a table with a CSV download option
datatable(
  avglog_n, 
  extensions = 'Buttons', 
  options = list(
    dom = 'Bfrtip',
    buttons = list('csv'),
    filter = "top"  # Ensures a search box appears
  )
)
```

### Top 20 Markers (pctdiff)

```{r,results='asis', message=FALSE, warning=FALSE}
datatable(
  pctdiff_n, 
  extensions = 'Buttons', 
  options = list(
    dom = 'Bfrtip',
    buttons = list('csv'),
    filter = "top"
  )
)
```

### Top 20 Markers (pct.ratio)

```{r,results='asis', message=FALSE, warning=FALSE}
datatable(
  pctratio_n, 
  extensions = 'Buttons', 
  options = list(
    dom = 'Bfrtip',
    buttons = list('csv', 'search')
  )
)
```


## All markers

While downloading tables please make sure to change filename, otherwise it might overwrite if downloading multiple tables. Make sure to include ".csv" after changing filename to retain format.

```{r,results='asis', message=FALSE, warning=FALSE}
# Create a table with a CSV download option
datatable(
  data.markers, 
  extensions = 'Buttons', 
  options = list(
    dom = 'Bfrtip',
    buttons = list('csv')
  )
)
```

```{r, results = FALSE, eval = FALSE, message=FALSE, warning=FALSE}
# Code block to find condition wise markers for each subcluster of subsetted data

# Read Seurat object
subset_data <- readRDS("../RDS/subset_umap_10_clusters.RDS")

# Ensure RNA assay is set
DefaultAssay(subset_data) <- "RNA"
Idents(subset_data) <- "seurat_clusters"

# Store subcluster identities (backup)
subset_data$subcluster <- Idents(subset_data)

# Set condition as active identity
Idents(subset_data) <- "condition"

subset_data <- JoinLayers(subset_data)

# Create an empty list to store results
markers_list <- list()

# Loop over each subcluster
for (cluster in unique(subset_data$subcluster)) {
  # Subset data for the current subcluster
   message("Processing Cluster: ", cluster)
   subcluster_data <- subset(subset_data, subset = subcluster == cluster)
 
  # Set condition as active identity
  Idents(subcluster_data) <- "condition"

  # Find all markers between Condition and Control within this subcluster
  markers <- FindAllMarkers(subcluster_data, only.pos = FALSE, min.pct = 0.25, logfc.threshold = 0.25)
   markers$p_val_adj <- p.adjust(markers$p_val, method = "fdr")

  # Store results with subcluster name
   markers_list[[paste0("subset_subcluster_", cluster)]] <- markers
}

# Save results to CSV
for (name in names(markers_list)) {
  write.csv(markers_list[[name]], file=paste0("../csvs/subset/condition/", name, "_Condition_vs_Control.csv"))
}
```


```{r, results = FALSE, message=FALSE, warning=FALSE}
# Calculating significant number of DEGs in each subcluster based on above generated files. Using p_val_adj < 0.05 and abs(avg_log2FC) > 1 to count a gene as significant

# Get list of all subcluster DEG files
subcluster_files <- list.files("../csvs/subset/condition/", pattern = "_Condition_vs_Control.csv", full.names = TRUE)

# Initialize empty data frame to store counts
deg_summary <- data.frame(Subcluster = character(), Num_Sig_DEGs = numeric(), stringsAsFactors = FALSE)

# Loop through each file and count significant DEGs
for (file in subcluster_files) {
  # Read data
  deg_data <- read.csv(file)
  
  # Filter for significant DEGs
  sig_degs <- deg_data %>% filter(p_val_adj < 0.05 & abs(avg_log2FC) > 1) %>%
  filter(!grepl("^MT", gene) & !grepl("^RB[SL]", gene))  # Exclude mitochondrial and ribosomal genes
  
  # Extract subcluster name from file name
  subcluster_name <- gsub("../csvs/subset/condition/|_Condition_vs_Control.csv", "", file)
  
  # Store result
  deg_summary <- rbind(deg_summary, data.frame(Subcluster = subcluster_name, Num_Sig_DEGs = nrow(sig_degs)))
}

# Save summary to CSV
write.csv(deg_summary, "../csvs/DEG_Summary.csv", row.names = FALSE)
```

# Number of significant DEGs (Condition vs Control) in each subcluster {.tabset}

- p_val_adj < 0.05, and abs(log2fc) > 1

- Not counting mitochondrial and ribosomal genes. They are still present in the data, but not being counted as DEGs.


*p_val_adj was calculated using FDR instead of default BH which causes inflated values.

```{r, results='asis', message=FALSE, warning=FALSE}
# Print summary
paged_table(deg_summary)
```

# Volcano plot for DEGs for all subclusters {.tabset}

Not plotting mitochondrial and ribosomal genes. Present in the data but not including them as genes of interest.

Highlighting genes which are: abs(log2fc) > 0.25, p_val_adj < 0.05

```{r,results='asis', message=FALSE, warning=FALSE}
# Code generates volcano plots for DEGs of interest. Input here is the previously generated condition files for each subcluster

library(ggplot2)
library(ggrepel)

# List of CSV files for each subcluster
csv_files <-  list.files("../csvs/subset/condition/", pattern = "_Condition_vs_Control.csv", full.names = TRUE)

# Function to generate volcano plot
volcanoplot <- function(de, subcluster_name) {
  
  # Filter out mitochondrial and ribosomal genes
  de <- de %>% filter(!grepl("^MT", gene) & !grepl("^RB[SL]", gene))

  # Subset input data for only condition
  de<- subset(de, cluster=="Condition")
  
  de$diffexpressed <- "NO"
  de$diffexpressed[de$avg_log2FC < -0.25 & de$p_val < 0.05] <- "DOWN"
  de$diffexpressed[de$avg_log2FC > 0.25 & de$p_val < 0.05] <- "UP"

  subset_df <- de[complete.cases(de[, "gene"]), ]

  c1 <- subset_df[subset_df[, "avg_log2FC"] > 0 & subset_df[, "p_val"] < 0.05, ]
  c2 <- subset_df[subset_df[, "avg_log2FC"] < 0 & subset_df[, "p_val"] < 0.05, ]
  c1_byavg_log2FC <- c1[order(c1$avg_log2FC, decreasing = TRUE), ]
  c2_byavg_log2FC <- c2[order(c2$avg_log2FC, decreasing = FALSE), ]
  c1_top10 <- head(c1_byavg_log2FC, 10)$gene
  c2_top10 <- head(c2_byavg_log2FC, 10)$gene
  labeled_proteins <- c(c1_top10, c2_top10)
  de$delabel <- ifelse(de$gene %in% labeled_proteins & de$diffexpressed %in% c("UP", "DOWN"), de$X, NA)

  ggplot(data = de, aes(x = avg_log2FC, y = -log10(p_val), col = diffexpressed, label = delabel)) +
    geom_vline(xintercept = c(-0.25, 0.25), col = "gray", linetype = 'dashed') +
    geom_hline(yintercept = -log10(0.05), col = "gray", linetype = 'dashed') + 
    geom_point(data = de %>% filter(diffexpressed == "DOWN"), color = "#00AFBB", size = 0.8, alpha = 0.5) +
    geom_point(data = de %>% filter(diffexpressed == "UP"), color = "#bb0c00", size = 0.8, alpha = 0.5) +
    geom_point(data = de %>% filter(diffexpressed == "NO"), color = "gray", size = 0.8, alpha = 0.5) +
    geom_text_repel(min.segment.length = 0, size = 2, max.overlaps = Inf, show.legend = FALSE) +
    scale_color_manual(values = c("DOWN" = "black", "UP" = "black")) +
    guides(color = guide_legend(title = NULL)) +
    labs(x = expression("Upregulated in CONTROL <- avg log"[2]*"FC -> Upregulated in Condition"), 
         y = expression("-log"[10]*"p-value"),
         title = paste0("DEGs in ", subcluster_name, " Condition vs CONTROL")) +
    theme_bw() +
    theme(plot.title = element_text(hjust = 0.5), axis.text = element_text(size = 16))
}

# Initialize list to store volcano plots
volcano_plots <- list()

# Loop through each CSV file, generate plots, and store them
for (file in csv_files) {
  if (grepl("subset_subcluster_\\d+_Condition_vs_Control.csv", file)) {
    subcluster_name <- gsub(".*subset_subcluster_(\\d+)_Condition_vs_Control.csv", "Subcluster \\1", file)
  } else if (grepl("subset_all_markers_condition.csv", file)) {
    subcluster_name <- "Subset (Overall)"
  } else {
    subcluster_name <- "Unknown"
  }
  
  de <- read.csv(file)
  
  # Store the plot in the list
  volcano_plots[[subcluster_name]] <- volcanoplot(de, subcluster_name)
}

# Print all volcano plots
for (name in names(volcano_plots)) {
  cat("##", name, "\n")
  print(volcano_plots[[name]])
  cat("\n\n")
}

```

