---
title: "CellChat"
author: "Saumya"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_depth: 3
    toc_float: true
    theme: cosmo
#    number_sections: true # Adds numbers to headers
    css: styles.css
    code_folding: hide
---

```{r, results = FALSE, message=FALSE, warning=FALSE}
# Load packages
library(dplyr)
library(Seurat)
library(patchwork)
library(ggplot2)
library(cowplot)
library(rmarkdown)
library(DT)
library(CellChat)
library(biomaRt)
library(ggalluvial)
library(extrafont)
library('NMF')
library('ggalluvial')
library(ComplexHeatmap)
```

```{r, results = FALSE, message=FALSE, warning=FALSE}
sampleNames <- c("ABC", "XYZ", "LMN")
data_dir <- ("../../raw_h5")
projectName <- "Seurat"
h5_string <- "_filtered_feature_bc_matrix.h5"
```

<!-- Tutorial: https://zhilongjia.github.io/scRNA/cellchat.html -->

```{r, eval = FALSE, results = FALSE, message=FALSE, warning=FALSE}
data.combined <- readRDS("../RDS/data_umap_29_clusters.RDS")
data.combined <- JoinLayers(data.combined)
```

```{r, eval = FALSE, results = FALSE, message=FALSE, warning=FALSE}
DefaultAssay(data.combined) <- "RNA"

# createCellChat() requires metadata column to have a "samples" column
data.combined$samples <- data.combined$sample_names

# Subset seurat object to condition of interest
data_control <- subset(data.combined, condition == "Control")

# Create cellchat object based on annotated clusters
cellchat <- createCellChat(object = data_control, group.by = "cluster_label")

# Load human cellchatDB and store in created cellchat object
CellChatDB <- CellChatDB.human
cellchat@DB <- CellChatDB

# Subset cellchat object to retain only genes present in DB as either ligands or receptors (hence have known interactions)
cellchat <- subsetData(cellchat)

# Identify overexpressesd genes
cellchat <- identifyOverExpressedGenes(cellchat)

# Identify overxpresed interactions based on overexpresse genes
cellchat <- identifyOverExpressedInteractions(cellchat)

# Commute probabilities of interactions and filter if interaction is in at least 10 cells in the group. Creates $prob and $pval columns in @net
cellchat <- computeCommunProb(cellchat)
cellchat <- filterCommunication(cellchat, min.cells = 10)

# Commute pathway probabilties. Add $prob and $pval in @netP, number of variables = number of pathways
cellchat <- computeCommunProbPathway(cellchat) #use projected data, raw.use = TRUE, uses non-projected data

# Aggregate interactions for each group/cluster, Creates $count and $weight columns for each group in @net with respect to other cell types/groups
cellchat <- aggregateNet(cellchat)

# Clustering based on functional similarity
cellchat <- computeNetSimilarity(cellchat, type = "functional")
cellchat <- netEmbedding(cellchat, type = "functional")
cellchat <- netClustering(cellchat, type = "functional",do.parallel = FALSE)

# Clustering based on structural similarity
cellchat <- computeNetSimilarity(cellchat, type = "structural")
cellchat <- netEmbedding(cellchat, type = "structural")
cellchat <- netClustering(cellchat, type = "structural", do.parallel = FALSE)

# For key signaling roles
cellchat <- netAnalysis_computeCentrality(cellchat, slot.name = "netP")

# For river plot
# Plot Cophenetic and Silhouette values to pick nPatterns.
selectK(cellchat, pattern = "incoming")

nPatterns = 3
cellchat <- identifyCommunicationPatterns(cellchat, pattern = "incoming", k = nPatterns)
netAnalysis_river(cellchat, pattern = "incoming")

## data frame cellchat information as a dataframe
df.net <- subsetCommunication(cellchat)
write.csv(df.net, "../csvs/cellchat_df_control.csv")

saveRDS(cellchat, "../RDS/cellchat_control.RDS")
```


```{r, results = FALSE, message=FALSE, warning=FALSE}
cellchat <- readRDS('../RDS/cellchat_control.RDS')
```

# All celltype information {.tabset}

```{r, results='asis',  message=FALSE, warning=FALSE}
cat("##", "Number of interactions", "\n")
netVisual_circle(cellchat@net$count, vertex.weight = as.numeric(table(cellchat@idents)),
                 weight.scale = TRUE, label.edge = FALSE, title.name = "Number of interactions")
cat("\n \n")

cat("##", "Weight of interactions", "\n")
netVisual_circle(cellchat@net$weight, vertex.weight = as.numeric(table(cellchat@idents)),
                 weight.scale = TRUE, label.edge= FALSE, title.name = "Interaction weights/strength")
cat("\n \n")
```


# Individual group interactions {.tabset}
```{r, results='asis', message=FALSE, warning=FALSE}
mat <- cellchat@net$weight
par(mfrow = c(1,1), xpd=TRUE)
for (i in 1:nrow(mat)) {
  cat("##", rownames(mat)[i], " \n")
  mat2 <- matrix(0, nrow = nrow(mat), ncol = ncol(mat), dimnames = dimnames(mat))
  mat2[i, ] <- mat[i, ]
  netVisual_circle(mat2, vertex.weight = as.numeric(table(cellchat@idents)), weight.scale = T, edge.weight.max = max(mat), title.name = rownames(mat)[i])
  cat(" \n \n")
}
```

```{r, results = FALSE, message=FALSE, warning=FALSE}
# To list all pathways
cellchat@netP[["pathways"]]

# Extract enriched LR pairs
extractEnrichedLR(cellchat, signaling = c(cellchat@netP[["pathways"]]), geneLR.return = TRUE)

# Visualise contribution of each LR pairs to the communication network
netAnalysis_contribution (cellchat, signaling = c(cellchat@netP[["pathwyas"]]), title = "Contribution of each LR pairs")

# Visualise specific pathways
netAnalysis_contribution (cellchat, signaling = c(cellchat@netP[["pathwyas"]][1:5]), title = "MIF, GALECTIN, CXCL, IL2, COMPLEMENT")

# Extract specific enriched pathway and visualise it
extractEnrichedLR(cellchat, signaling = c("TNF"), geneLR.return = FALSE)
netAnalysis_contribution (cellchat, signaling = c("TNF"))

# Visualise aggregate pathway information for a specific LR
netVisual_aggregate(cellchat, signaling = "TNF", layout = "circle")
#Visualise a circle plot for each pathway instead of aggregating for that ligand/receptor
netVisual_individual(cellchat, signaling = "TNF", layout = "circle")
# You can also choose a specific pathway to be plotted instead of plotting all
netVisual_individual(cellchat, signaling = "TNF", pairLR.use = "TNF_TNFRSF1A", layout = "circle")

# Chord diagram plots genes in pathway interaction. One side is the ligand and the other receptor
netVisual_chord_gene(cellchat, signaling = "TNF")

# Chord diagram plots by groups of cell types. E.g, myeloid, endothelial, etc.
group.cellType <- c(rep("FIB", 4), rep("DC", 4), rep("TC", 4))
names(group.cellType) <- levels(cellchat@idents)
par(mfrow = c(1, 1), xpd=TRUE)
par(cex = 0.5)
netVisual_chord_cell(cellchat, signaling = "TNF",
    group = group.cellType,
    title.name = paste0("TNF_", "signaling network"))

# Chord diagram: show LR pairs associated with certain signaling pathways. E.g, stroma to ILC
netVisual_chord_gene(cellchat, sources.use = c(1,2,3,4), targets.use = 8,
                     signaling = c("CCL","CXCL"),legend.pos.x = 8)

# Violin plot
plotGeneExpression(cellchat, signaling = "TNF")

# bubble plot
# bubble plot: show all LR pairs from source to target cell groups
netVisual_bubble(cellchat, sources.use = 4, targets.use = c(5:12),
                 remove.isolate = FALSE)

# bubble plot: show LR pairs associated with certain signaling pathways
netVisual_bubble(cellchat, sources.use = 4, targets.use = c(5:12),
                 signaling = c("CCL","CXCL"), remove.isolate = FALSE)

# Scatter plot to visualize aggregated communication networks for each cell type
netAnalysis_signalingRole_scatter(cellchat) # all signaling pathways

# Scatter plot to Visualize selected communication networks
netAnalysis_signalingRole_scatter(cellchat, signaling = "TNF")
netAnalysis_signalingRole_scatter(cellchat, signaling = c("CXCL", "CCL"))

# Heatmap to visualize dominant cell types for each signaling pathway
netAnalysis_signalingRole_heatmap(cellchat, pattern = "outgoing", height = 11)
netAnalysis_signalingRole_heatmap(cellchat, pattern = "incoming", height = 11)

# Visualize selected outgoing/incoming signals and contributing cell types
netAnalysis_signalingRole_heatmap(cellchat, pattern = "outgoing",
                                      signaling = c("CXCL", "CCL"))
netAnalysis_signalingRole_heatmap(cellchat, pattern = "incoming",
                                      signaling = c("CXCL", "CCL"))

# Heatmap to visualize major signaling roles of different cell groups
netAnalysis_signalingRole_network(cellchat, signaling = "TNF", width = 10,
                                  height = 5, font.size = 10)
```

# Find important signal pathway {.tabset}


```{r, results='asis', message=FALSE, warning=FALSE}
cat("##", "Functional", " \n")
netVisual_embeddingZoomIn(cellchat, type = "functional", label.size = 3.5, nCol = 1) + ggtitle ("Structural classification of signaling networks")
cat(" \n \n")

cat("##", "Structual", " \n")
 netVisual_embeddingZoomIn(cellchat, type = "structural", label.size = 3, nCol = 1) + ggtitle ("Structural classification of signaling networks")
cat(" \n \n")
```

```{r, results='asis', message=FALSE, warning=FALSE}

pathways.show <- c("CXCL") 
vertex.receiver = c(1:10) 

# Hierarchical plot
cat("##", "Hierarchical", " \n")
netVisual_aggregate(cellchat, signaling = pathways.show, vertex.receiver = vertex.receiver, layout = "hierarchy", pt.title = 14, title.space = 4, vertex.label.cex = 0.8)
par(mfrow=c(1,1))
cat(" \n \n")

# Circular plot
cat("##", "Circular", " \n")
netVisual_aggregate(cellchat, signaling = pathways.show, layout = "circle")
par(mfrow=c(1,1))
cat(" \n \n")

# Chord diagram
cat("##", "Chord", " \n")
netVisual_aggregate(cellchat, signaling = pathways.show, layout = "chord")
par(mfrow=c(1,1))
cat(" \n \n")

# Heatmap
cat("##", "Heatmap", " \n")
netVisual_heatmap(cellchat, signaling = pathways.show, color.heatmap = "Reds")
cat(" \n \n")
```

<!-- CellChat Comparative Analysis. Requires at least 2 cellchat condition RDS for comparisons -->


```{r, eval = FALSE, results = FALSE, message=FALSE, warning=FALSE}
cellchat_condition <- readRDS("../RDS/cellchat_condition.RDS")
cellchat_control <- readRDS("../RDS/cellchat_control.RDS")
 
object.list <- list(Condition = cellchat_condition, NV = cellchat_control)

# Check list
names(object.list)
 
cellchat <- mergeCellChat(object.list, add.names = names(object.list))

# Save merged cellchat object
saveRDS(cellchat, "../RDS/cellchat_merged.RDS")

# Save object list for later use
saveRDS(object.list, "../RDS/cellchat_merged_object_list.RDS")
```


```{r, results = FALSE, message=FALSE, warning=FALSE}
cellchat <- readRDS("../RDS/cellchat_merged.RDS")
object.list <- readRDS("../RDS/cellchat_merged_object_list.RDS")
```

# 1. Overall Pathway Comparison between Condition and NV {.tabset}

Colour coded based on enrichment + whether that difference is statistically significant. Black means equal in both.

```{r, results='asis', message=FALSE, warning=FALSE}
# 1. Compare the overall information flow of each signaling pathway
cat("##", "Overall Pathway Flow", "\n")
rankNet(cellchat, mode = "comparison", stacked = T, do.stat = TRUE) 
#rankNet(cellchat, mode = "comparison", stacked = F, do.stat = TRUE) To plot red/blue for each pathway
cat("\n \n")

#rankNet(cellchat, mode = "comparison", stacked = T, do.stat = TRUE, sources.use = "PC", title = "PC")
```
# 2. Total Number of Interactions and Interaction Strength {.tabset}

```{r, results='asis', message=FALSE, warning=FALSE}
# 2. Compare the total number of interactions and interaction strength
cat("##", "Total Count", "\n")
compareInteractions(cellchat, show.legend = F, group = c(1,2), measure = "count")
cat("\n \n")

cat("##", "Total Weight", "\n")
compareInteractions(cellchat, show.legend = F, group = c(1,2), measure = "weight")
cat("\n \n")
```

# 3. Outgoing/Incoming interaction strength for all cell types {.tabset}

```{r, results='asis', message=FALSE, warning=FALSE}
###### scatter plot 
# 1.Compare outgoing/incoming interaction strength for all the cell types
count.sum <- sapply(object.list, function(x) {rowSums(x@net$count) + 
    colSums(x@net$count)-diag(x@net$count)})
weight.MinMax <- c(min(count.sum), max(count.sum)) # control the dot size 
gg <- list()
for (i in 1:length(object.list)) {
  gg[[i]] <- netAnalysis_signalingRole_scatter(object.list[[i]], 
  title = names(object.list)[i], weight.MinMax = weight.MinMax)
}

patchwork::wrap_plots(plots = gg)

# cat("##", "Condition", "\n")
# gg[1]
# cat("\n \n")
# 
# cat("##", "NV", "\n")
# gg[2]
# cat("\n \n")
```


#  4. Signaling changes associated with one cell group {.tabset}

Scatter plot shows the relative changes in overall signaling strength for each pathway between two groups. The axes represent:
X-axis: Information flow (i.e., total communication probability) in Group 1 (Condition)
Y-axis: Information flow in Group 2 (NV)
The size of the dot usually reflects the overall communication strength (sum of probabilities).

Positive values = upregulated, negative values = downregulated
Can include/exclude specific pathways if need be.

No set cutoff threshold but:

| Value  | Range | Interpretation |
|--|--|--|
|< 0.01	|Very weak | Noise/background
| 0.01–0.05 | Weak to moderate |Worth considering if specific or known pathways
| 0.05–0.1 | Moderate |If consistent across samples
| 0.1 |Strong  |More biologically relevant
|0.2 or 0.3|Very strong|Important biologically

```{r, results='asis', message=FALSE, warning=FALSE}
# 2. identify signalling changes associated with one cell group 

cat("##", "Subset", "\n")
netAnalysis_signalingChanges_scatter(cellchat, idents.use = "Subset")
cat("\n \n")
```

# 5. Circle Plots: Number of interactions between any two cell populations {.tabset}

Can pick specific cell types

```{r, results='asis',  message=FALSE, warning=FALSE}
# 1. show the number of interactions between any two cell populations 
# compute the maximum number of cells and the maximum number of interactions 
weight.max <- getMaxWeight(object.list, attribute = c("idents","count"))

par(mfrow = c(1,2), xpd=TRUE)
cat("##", "All Pathways", "\n")
for (i in 1:length(object.list)) {
  netVisual_circle(object.list[[i]]@net$count, weight.scale = T, label.edge= F,
  edge.weight.max = weight.max[2], edge.width.max = 12, arrow.size = 0.1,
  title.name = paste0("Number of interactions - ", names(object.list)[i]))
}
cat("\n \n")


# 2. selected pathway
weight.max <- getMaxWeight(object.list, slot.name = c("netP"), 
                           attribute =c("CXCL"))

par(mfrow = c(1,2), xpd=TRUE)
cat("##", "Specific Pathway", "\n")
for (i in 1:length(object.list)) {
  netVisual_aggregate(object.list[[i]], signaling = c("CXCL"), layout = "circle",
  edge.weight.max = weight.max[1], edge.width.max = 10, arrow.size = 0.05, 
  signaling.name = paste("CXCL", names(object.list)[i]))
}
cat("\n \n")
```

# 5. Circle Plot - Differential Expression {.tabset}

Condition vs NV

Red arrows - upregulation
Blue arrows - downregulation

Can pick specific cell types

```{r, results='asis', message=FALSE, warning=FALSE}
# 3. Show differential number of interactions or interaction strength among 
# different cell populations, red(increased signaling)/blue(decreased signaling)

par(mfrow = c(1,2), xpd=TRUE)
netVisual_diffInteraction(cellchat, comparison = c(1, 2), measure = "count", 
                          weight.scale = T, arrow.size = 0.1)
netVisual_diffInteraction(cellchat, comparison = c(1, 2), measure = "weight", 
                          weight.scale = T, arrow.size = 0.1)


# 4. simplify the complicated network to the cell type level - pick cell types of interest.
group.cellType <- c(rep("FIB", 4), rep("DC", 4), rep("TC", 4))
group.cellType <- factor(group.cellType, levels = c("FIB", "DC", "TC"))
object.list <- lapply(object.list, function(x) {
                               mergeInteractions(x, group.cellType)})
cellchat <- mergeCellChat(object.list, add.names = names(object.list))
 
 weight.max <- getMaxWeight(object.list, slot.name = c("idents", "net", "net"), 
                            attribute = c("idents", "count", "count.merged"))
 
# show the number of interactions or interaction strength.
par(mfrow = c(1,2), xpd=TRUE)
for (i in 1:length(object.list)) {
     netVisual_circle(object.list[[i]]@net$count.merged, weight.scale = T, 
     label.edge= T, edge.weight.max = weight.max[3], edge.width.max = 12, 
     arrow.size = 0.1,
     title.name = paste0("Number of interactions - ", names(object.list)[i]))
}
```

# 6. Heatmaps {.tabset}

In differential heatmap, red = upregulated, blue = downregulated

```{r, results='asis',  message=FALSE, warning=FALSE, fig.height= 8}

# 1. Compare outgoing/incoming signaling associated with each cell population
# combining all the identified signaling pathways from different datasets 
all_pathways <- union(object.list[[1]]@netP$pathways, 
                       object.list[[2]]@netP$pathways)

ht1 = netAnalysis_signalingRole_heatmap(object.list[[1]], pattern = "all", 
      signaling = all_pathways, title = names(object.list)[1],  
      width = 5, height = 11, color.heatmap = "OrRd")
ht2 = netAnalysis_signalingRole_heatmap(object.list[[2]], pattern = "all", 
      signaling = all_pathways, title = names(object.list)[2], 
      width = 5, height = 11, color.heatmap = "OrRd")
cat("##", "Overall Heatmap", " \n")
draw(ht1 + ht2, ht_gap = unit(0.5, "cm"))
cat(" \n \n")

ht3 = netAnalysis_signalingRole_heatmap(object.list[[1]], pattern ="outgoing", 
      signaling = all_pathways, title = names(object.list)[1], 
      width = 5, height = 11)
ht4 = netAnalysis_signalingRole_heatmap(object.list[[2]], pattern ="outgoing", 
      signaling = all_pathways, title = names(object.list)[2], 
      width = 5, height = 11)
cat("##", "Outgoing Heatmap", " \n")
draw(ht3 + ht4, ht_gap = unit(0.5, "cm"))
cat(" \n \n")


ht5 = netAnalysis_signalingRole_heatmap(object.list[[1]], pattern = "incoming", 
      signaling = all_pathways, title = names(object.list)[1], 
      width = 5, height = 11, color.heatmap = "GnBu")
ht6 = netAnalysis_signalingRole_heatmap(object.list[[2]], pattern ="incoming",
      signaling = all_pathways, title = names(object.list)[2], 
      width = 5, height = 11, color.heatmap = "GnBu")
cat("##", "Incoming Heatmap", " \n")
draw(ht5 + ht6, ht_gap = unit(0.5, "cm"))
cat(" \n \n")

# 2. selected pathways
par(mfrow = c(1,2), xpd=TRUE)
ht <- list()
for (i in 1:length(object.list)) {
  ht[[i]] <- netVisual_heatmap(object.list[[i]], signaling = c("CXCL"), 
  title.name = paste("CXCL", "signaling ",names(object.list)[i]),
  color.heatmap = "Reds")
}
cat("##", "CXCL pathway", " \n")
ComplexHeatmap::draw(ht[[1]] + ht[[2]], ht_gap = unit(0.5, "cm"))
cat(" \n \n")

# 3. Differential Interactions
gg1 <- netVisual_heatmap(cellchat, comparison = c(1, 2), measure = "count")
gg2 <- netVisual_heatmap(cellchat, comparison = c(1, 2), measure = "weight")
cat("##", "Differential Heatmap", " \n")
gg1 + gg2
cat(" \n \n")
```
# 7. Bubble Plot {.tabset}

```{r, results='asis', message=FALSE, warning=FALSE}
# Chord diagram

# 1. compare communication probabilities mediated by ligand-receptor pairs from 
# selected sources and targets cell groups 
cat("##", "Overall Bubbleplot", " \n")
netVisual_bubble(cellchat, sources.use = 9, targets.use = c(10:19),  
                 comparison = c(1, 2), angle.x = 45)
cat(" \n \n")

# 2. identify the up-regulated ligand-receptor pairs
cat("##", "Top Upregulated LRs in Condition", " \n")
netVisual_bubble(cellchat, sources.use = 9, targets.use = c(10:19),
                 comparison = c(1, 2), max.dataset = 1,
                 title.name = "Increased signaling in Condition", angle.x = 45,
                 remove.isolate = T)
cat(" \n \n")

# 2. identify the down-regulated ligand-receptor pairs
cat("##", "Top Downregulated LRs in Condition", " \n")
netVisual_bubble(cellchat, sources.use = 9, targets.use = c(10:19),  
                 comparison = c(1, 2), max.dataset = 2, 
                 title.name = "Decreased signaling in HIV", angle.x = 45, 
                 remove.isolate = T) 
cat(" \n \n")
```

# 8. Violin Plot {.tabset}
For a specific pathway
```{r, results='asis',  message=FALSE, warning=FALSE}
cat("##", "CXCL Pathway", " \n")
plotGeneExpression(cellchat, signaling = "CXCL", split.by = "condition", 
                   colors.ggplot = T)
cat(" \n \n")

```




