---
title: "Trajectory Analysis"
author: "Saumya"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_depth: 3
    toc_float: true
    theme: cosmo
#    number_sections: true # Adds numbers to headers
    css: styles.css
    code_folding: hide
---

```{r, echo=FALSE}
colorize <- function(text, color) {
  sprintf('<span style="color:%s">%s</span>', color, text)
}
```

<!-- Information of Used Library -->
### Information of Used R Libraries

The following R libraries were used to process the data and produce this report.

dplyr: `r packageVersion('dplyr')`, 

Seurat: `r packageVersion('Seurat')`, 

patchwork: `r packageVersion('patchwork')`, 

ggplot2: `r packageVersion('ggplot2')`, 

cowplot: `r packageVersion('cowplot')`,

Monocle3 : `r packageVersion('monocle3')`,

SeuratWrappers: `r packageVersion('SeuratWrappers')`

# Trajectory/Pseudotime Analysis using Monocle3


```{r, eval=FALSE,echo=FALSE, message=FALSE, warning=FALSE}
# Load the Seurat object from RDS file
seurat_object <- readRDS("seurat_obj.RDS")

# Join layers if the object has multiple layers (e.g., from integration) for Seurat V5
seurat_object <- JoinLayers(seurat_object)

# Create a working copy of the Seurat object
my.so <- seurat_object

# Project PCA dimensions to the whole dataset for consistency
my.so <- ProjectDim(my.so, reduction = "pca")
```

```{r, eval=FALSE,echo=FALSE, message=FALSE, warning=FALSE}
# Create an expression matrix from the Seurat object
# Extract raw count data from the RNA assay
expression_matrix <- GetAssayData(my.so, assay = "RNA", layer = "counts")

# Get cell metadata (cell annotations, cluster info, etc.)
cell_metadata <- my.so@meta.data
cell_metadata <- cell_metadata[colnames(expression_matrix), ]

# Verify that cell identifiers match between expression matrix and metadata
# This is crucial for Monocle3 to work properly

if (isTRUE(all.equal(colnames(expression_matrix), rownames(cell_metadata)))) {
  print("Cell identifiers match")
} else {
  print(sprintf("Cell identifier mismatch - %i cells in expression matrix, %i cells in metadata",
                ncol(expression_matrix), nrow(cell_metadata)))
  print("If the counts are equal, sort differences will throw this error")
}
```

```{r,eval=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# Create gene annotations for Monocle3
# Extract gene names from the Seurat object and create a data frame

gene_annotation <- data.frame(gene_short_name = rownames(my.so@assays$RNA), row.names = rownames(my.so@assays$RNA))

# Verify that gene identifiers match between expression matrix and gene annotations
# This ensures all genes are properly labeled in Monocle3
if (all.equal(rownames(expression_matrix), rownames(gene_annotation))) {
  print(sprintf("Gene identifiers all match"))
} else {
  print(sprintf("Gene identifier mismatch - %i genes in expression matrix, %i gene in gene annotation",
                nrow(expression_matrix), nrow(gene_annotation)))
  print("If the counts are equal, sort differences will throw this error")
}
```

```{r, eval=FALSE,echo=FALSE, message=FALSE, warning=FALSE}
# Create a Monocle3 cell data set (CDS) object from Seurat data
# This converts Seurat object format to Monocle3 format for trajectory analysis

cds <- new_cell_data_set(expression_matrix,
                            cell_metadata = cell_metadata,
                            gene_metadata = gene_annotation)
```


```{r, eval=FALSE,echo=FALSE, message=FALSE, warning=FALSE}
# Transfer Seurat embeddings and cluster info to Monocle3 object
# This ensures Monocle3 uses the same PCA and UMAP coordinates as Seurat

reducedDim(cds, type = "PCA") <- my.so@reductions$pca@cell.embeddings 
cds@reduce_dim_aux$prop_var_expl <- my.so@reductions$pca@stdev

# Optional: Plot PCA variance explained to verify the transfer worked correctly
# plot_pc_variance_explained(cds)
```

```{r, eval=FALSE,echo=FALSE, message=FALSE, warning=FALSE}
# Transfer UMAP coordinates from Seurat to Monocle3
# This ensures both objects use the same UMAP visualization

cds@int_colData@listData$reducedDims$UMAP <- my.so@reductions$umap@cell.embeddings

# Plot cells using Monocle3's plotting function
# This should show the same UMAP as Seurat

plot_cells(cds)
```

```{r, eval=FALSE,echo=FALSE, message=FALSE, warning=FALSE}
# Transfer Seurat cluster information to Monocle3 object
# This ensures Monocle3 uses the same clustering as Seurat

cds@clusters$UMAP_so$clusters <- my.so@meta.data$seurat_clusters

# Perform clustering in Monocle3 using the transferred cluster information
# This creates a consistent clustering between Seurat and Monocle3

cds <- cluster_cells(cds, reduction_method = "UMAP", partitioning_method = "seurat_clusters")
```

```{r, eval=FALSE,echo=FALSE, message=FALSE, warning=FALSE}
# Save the Monocle3 cell data set object for later use
# This preserves all the clustering and embedding information
saveRDS(cds, file = "trajectory_cds.RDS")
```

## Seurat and Monocle3 clusters {.tabset}
```{r, results='asis', echo=FALSE, message=FALSE, warning=FALSE}
# Load the Seurat object for visualization
my.so <- readRDS("seurat_obj.RDS")

# Load the Monocle3 cell data set object
cds <- readRDS("trajectory_cds.RDS")

# Plot clusters for comparison between Seurat and Monocle3 clustering
cat('###',"Seurat Clusters",' \n')
# Create UMAP plot colored by Seurat clusters with labels
DimPlot(my.so, reduction = "umap", group.by = "seurat_clusters", label = TRUE, raster = FALSE)+ggtitle("Seurat Cluster")
cat('\n \n')
cat('###',"Monocle3 Clusters",' \n')
# Create UMAP plot colored by Monocle3 clusters without labels
plot_cells(cds, color_cells_by = "cluster", label_cell_groups=FALSE)+ggtitle("Monocle Cluster")
cat('\n \n')
```

```{r, eval=FALSE,echo=FALSE, message=FALSE, warning=FALSE}
# Learn the trajectory graph from the cell data set
# This constructs a graph representing cell state transitions
# The graph will be used to order cells along developmental trajectories

cds <- learn_graph(cds)
```

```{r, eval=FALSE,echo=FALSE, message=FALSE, warning=FALSE}
# Extract the Seurat cluster information from the Monocle3 object
# This transfers cluster labels from Seurat to Monocle3 for consistency

cds@clusters$seurat_clusters <- my.so@meta.data$seurat_clusters

# Identify the root cells for trajectory analysis
# Root cells define the starting point of the trajectory

root = c("0")
root_cells <- rownames(subset(colData(cds), seurat_clusters == root))

# Print cluster distribution and number of root cells for verification
print(table(cds@clusters$seurat_clusters))
print(length(root_cells))
```

```{r, eval=FALSE,echo=FALSE, message=FALSE, warning=FALSE}
# Order cells in the Monocle3 object using the identified root cells
# This assigns pseudotime values to each cell based on distance from root
# Root cells get pseudotime = 0, cells further along trajectory get higher values

cds <- order_cells(cds, root_cells = root_cells)
```

```{r, eval=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# Save the final Monocle3 object with trajectory information
# This includes the learned graph and is ready for pseudotime analysis

saveRDS(cds, file = "monocle3_cds.RDS")
```

## Monocle3 Pseudotime Analysis {.tabset}
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Plot cells with pseudotime values

cds <- readRDS("monocle3_cds.RDS")

# Create UMAP plot colored by pseudotime values
# Cells are colored based on their pseudotime (developmental progression)

seurat_pseudotime <- plot_cells(cds, reduction_method = "UMAP", cell_size = 0.5, color_cells_by = "pseudotime") + ggtitle("Monocle3 Pseudotime Analysis")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Plot cells colored by their cell type/cluster identity
# This shows how different cell types are distributed along the trajectory

celltype_pseudotime <- plot_cells(
  cds, 
  reduction_method = "UMAP", 
  cell_size = 0.5, 
  color_cells_by = "seurat_clusters", 
  label_cell_groups = FALSE
) + theme(legend.position = "right") + ggtitle("Pseudotime Analysis with Cell Type")
```

```{r, results='asis', echo=FALSE, message=FALSE, warning=FALSE}
# Display pseudotime analysis results in the report
cat('###',"Pseudotime",' \n')
# Show the pseudotime plot (cells colored by developmental progression)
print(seurat_pseudotime)
cat('\n \n')
cat('###',"Pseudotime with Cell Type",' \n')
# Show the cell type plot (cells colored by cluster identity)
print(celltype_pseudotime)
cat('\n \n')
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Save pseudotime plot as PDF file

pdf(paste("trajectory_analysis.pdf"), width=10, height=8, useDingbats = FALSE)
print(seurat_pseudotime)
print(celltype_pseudotime)
dev.off()
```